### 2017-12-08 article/article.tex
Add report

\documentclass[letterpaper]{article}
\usepackage{natbib,alifexi}

\title{TODO title}
\author{Cédric Simar, Antoine Passemiers, Stanislas Gueniffey \and Robin Petit \\
\mbox{}\\
Université Libre de Bruxelles \\
apassemi@ulb.ac.be}

\begin{document}
\maketitle

\begin{abstract}
  TODO
\end{abstract}

\section{Introduction}

TODO

\section{Methods}

TODO \citep{Mnih2015}

\section{Results}

TODO

\section{Discussion}

TODO

\section{Acknowledgements}

TODO

\footnotesize
\bibliographystyle{apalike}
\bibliography{article}

\end{document}

### 2017-12-13 article/article.tex
[Article] Abstract

@@
 \usepackage{natbib,alifexi}
+\usepackage{color}
 
-\title{TODO title}
-\author{Cédric Simar, Antoine Passemiers, Stanislas Gueniffey \and Robin Petit \\
+\newcommand\todo[1]{\textcolor{red}{TODO: #1}}
+
+\title{Human-level control through deep reinforcement learning}
+\author{C\'{e}dric Simar, Antoine Passemiers, Stanislas Gueniffey \and Robin Petit \\
 \mbox{}\\
-Université Libre de Bruxelles \\
+Universit\'{e} Libre de Bruxelles \\
 apassemi@ulb.ac.be}
@@
 \begin{abstract}
-  TODO
+
+  Reinforcement learning is a subarea of machine learning inspired by neuroscience, where learners have to select actions with
+  regards to the environement state and receive a reward accordingly. Each agent's objective is to maximize its cumulative reward across its whole lifetime,
+  subdivided into episodes. In the traditional framework of reinforcement learning and more specifically in Q-learning,
+  researchers usually deal with a small discrete space to represent the values of Q. This is no longer sufficient if one desires to describe the environment state
+  as an image of raw pixels. In consequence, we introduce deep Q-learning methods, which have been found to be really effective in mapping raw pixels
+  to abstract high-level features. In deep Q-learning, the estimation of Q-values for a given action are produced using deep learning neural networks,
+  which consist in stacks of many neural layers. These techniques have revealed themselves to be able to beat human experts at playing Atari games 
+  by learning only from visual features. Despite the current progress, learning to take complex decisions on the basis of high-dimensional visual data remains on
+  ongoing challenge.
+
 \end{abstract}
@@
 
-TODO
+  \todo{pour Robin}
 
@@
 
-TODO \citep{Mnih2015}
+  \todo{pour Robin} \citep{Mnih2015}
 
@@
 
-TODO
+  \todo{Plots avec les erreurs d'apprentissage (pas de cross-val)}
+
+  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop} \citep{wattenberg2016how}
 
@@
 
-TODO
+  \todo{Amélioration: Double dueling DQN} \citep{DBLP:journals/corr/WangFL15}
 
@@
 
-TODO
+  \todo{Remercier Lenaerts, Nowe, Google, OpenAI, Markov}
 

### 2017-12-14 article/article.tex
[Article] Add reference for ALE

@@
   as an image of raw pixels. In consequence, we introduce deep Q-learning methods, which have been found to be really effective in mapping raw pixels
-  to abstract high-level features. In deep Q-learning, the estimation of Q-values for a given action are produced using deep learning neural networks,
-  which consist in stacks of many neural layers. These techniques have revealed themselves to be able to beat human experts at playing Atari games 
+  to abstract high-level features. In deep Q-learning, the estimation of Q-values for a given action are produced using deep neural networks,
+  which consist inoof many neural layers stacked on top of each other. 
+  These techniques have revealed themselves to be able to beat human experts at playing Atari games 
   by learning only from visual features. Despite the current progress, learning to take complex decisions on the basis of high-dimensional visual data remains on
@@
   \todo{pour Robin} \citep{Mnih2015}
+
+  \todo{Citer ALE} \citep{bellemare13arcade}
 

### 2017-12-18 article/article.tex
[Article] Introduction: Q-value and CNN

@@
 \usepackage{color}
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
   to abstract high-level features. In deep Q-learning, the estimation of Q-values for a given action are produced using deep neural networks,
-  which consist inoof many neural layers stacked on top of each other. 
+  which consist of many neural layers stacked on top of each other. 
   These techniques have revealed themselves to be able to beat human experts at playing Atari games 
@@
 
-  \todo{pour Robin}
+We designed the agent in such a way that it selects actions according to its beliefs about future rewards, and its only objective is to maximize its total
+cumulative reward. More formally, this cumulative calculation starts from present time $t$ and is computed as follows:
+\[
+  Q^{*}(s, a) = \max_{\pi} \mathop{\mathbb{E}}[ r_t + \gamma r_{t+1} + \gamma^2 r_{t+2} + ... | s_t = s, a_t = a, \pi]
+\]
+where $\pi = P(a | s)$ is a behavior policy, $t$ is the current step, $r_t$ is the reward for step $t$, $a_t$ is the action selected at time $t$, 
+and $\gamma$ is the discount factor. Thus, we can described the Q-value as the present value of all future rewards at current step $t$,
+with respect to the environment state and the action selected at $t$.
+
+Because of the very high dimensionality of input visual features, it is not reasonable to consider using a table for storing all Q-values.
+Instead, one can design a learnable function that maps pixel values to a single row of this table, thus a vector of Q-values whose length is the size
+of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article \citep{Mnih2015}.
+A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers. One characteristic of CNNs is the
+presence of convolutional filters that maps pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
+allows the convolutional layer to capture the local information of the pixels, contrary to what dense (fully-connected) layers do. This procedure is greatly
+inspired by the notion of receptive field introduced by Hubel and Wiesel \citep{Hubel1962}.
+Let $Q(s, a; \theta_i)$ be the Q-value approximation function represented by the neural network, parameterized by its weights $\theta$ at time step $i$.
+
+Since we are handling visual features, we used screen frames as environment state representations. Each frame is a (210 x 160) color image sampled at
+a frequency of 60 Hz.
+
+  \todo{Antoine suite}
 
@@
 
-  \todo{pour Robin} \citep{Mnih2015}
+  \todo{pour Robin}
 

### 2017-12-18 agent.py
Just a start, doesn't learn shit yet, more job to be done in agent and dqn (plus structure the whole thing and have a proper main), the architecture of the NN should be more or less correct tough

@@ -1,0 +1,25 @@
+# -*- coding: utf-8 -*-
+# dqn.py
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
+
+from parameters import Parameters
+from environment import Environment
+from dqn import DQN
+
+import numpy as np
+import tensorflow as tf
+
+
+class Agent:
+    def __init__(self):
+        
+        self.environment = Environment()
+        self.dqn = DQN(self.environment.get_input())
+
+
+    def select_action(self, s):
+        # The agent uses its experience and expertise acquired
+        # through deep learning to make intelligent actions
+        return np.random.randint(0, self.n_actions, size=1)[0]
+
+Agent()

### 2017-12-18 dqn.py

@@ -1,20 +1,113 @@
-# -*- coding: utf-8 -*-
-# dqn.py
-# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-from utils import Params
-
-import numpy as np
 import tensorflow as tf
+from tf_decorator import *
+from parameters import Parameters
 
 
-class Agent:
-    def __init__(self, session, n_actions):
-        self.session = session # Tensorflow session
-        self.n_actions = n_actions
-        # TODO: init neural network
+class DQN:
 
-    def select_action(self, s):
-        # The agent uses its experience and expertise acquired
-        # through deep learning to make intelligent actions
-        return np.random.randint(0, self.n_actions, size=1)[0]
+    def __init__(self, image):
+
+        # receiving image placeholder
+        self.image = image
+
+        # initialize tensorflow graph
+        self.prediction
+        self.train
+        self.error
+
+
+    @define_scope
+    def prediction(self):
+        
+        """
+        [Article] The input to the neural network consists of an 84 x 84 x 4 image 
+        produced by the preprocessing map ϕ
+        """
+        # reshape input to 4d tensor [batch, height, width, channels]
+        input = tf.reshape(self.image, [-1, Parameters.image_height, Parameters.image_width, Parameters.m_recent_frames])
+
+        # convolutional layer 1
+        """
+        [Article] The first hidden layer convolves 32 filters of 8 x 8 with stride 4 with the
+        input image and applies a rectifier nonlinearity
+        """
+        W_conv1 = self.weight_variable([8, 8, Parameters.m_recent_frames, 32])
+        b_conv1 = self.bias_variable([32])
+        conv1 = tf.nn.conv2d(input, W_conv1, strides=[1, 4, 4, 1], padding='VALID') # would 'SAME' also work ?
+        h_conv1 = tf.nn.relu(conv1 + b_conv1)
+        
+        # output of conv 1 is of shape [-1 x 20 x 20 x 32]
+        
+        # convolutional layer 2
+        """
+        [Article] The second hidden layer convolves 64 filters of 4 x 4 with stride 2, 
+        again followed by a rectifier nonlinearity
+        """
[…]

### 2017-12-18 environment.py

@@ -1,0 +1,133 @@
+
+import gym
+import numpy as np
+import cv2
+from history import Frames_History
+from parameters import Parameters
+import utils
+import matplotlib.pyplot as plt
+
+# ϕ
+
+class Environment:
+
+    def __init__(self):
+
+        self.environment = gym.envs.make(Parameters.game)
+        self.new_game()
+
+        # let's have a look at the first cnn input
+        for screen in self.history.get():
+            plt.imshow(screen, cmap='gray')
+            plt.show()
+        
+
+    def new_game(self):
+        """
+        Create a new game and return the game variables
+        """
+
+        self._screen = self.environment.reset()
+        first_action = 0
+        self.take_action(first_action)
+        self.render()
+        self.reward = 0
+        self.terminal = False # True?
+
+        self.initialize_screens_history()
+
+        return(self.screen, first_action, self.reward, self.terminal)
+    
+
+    def initialize_screens_history(self):
+
+        """ initialize the history by pushing the first screen 4 (m_recent_frames) times """
+
+        self.history = Frames_History()
+        for _ in range(Parameters.m_recent_frames):
+            self.history.add_frame(self.screen)
+
+    
+    def add_current_screen_to_history(self):
+        self.history.add_frame(self.screen)
+
+    
+    def get_input(self):
+        return(self.history.get())
+
+
+    def select_random_action(self):
+        return(self.environment.action_space.sample())
+    
+
+    def select_smart_action(self):
+        """
+        Not smart yet
+        """
+        return(self.select_random_action())
+
+
+    def take_action(self, action):
+        self._previous_screen = self._screen
+        self._screen, self.reward, self.terminal, _ = self.environment.step(action)
+
+    
+    def act(self, action):
+
+        """
+        Take the action a certain number of times (Parameters.frame_skipping)
+        as described in the article
+        Return the environment state after taking the action x times
+        """
+
+        lives_before_action = self.lives
+        cumulated_reward = Parameters.no_reward
+
+        # frame skipping (see Parameters for more information)
+        skipped = 0 # break is evil
+        while(skipped < Parameters.frame_skipping and not self.terminal):
+            
+            self.take_action(action)
+            cumulated_reward += self.reward
+
+            ## TODO ?? Should we update screen history or not here ?? ##
+
+            """
+            [Article] For games where there is a life counter, the Atari
+            2600 emulator also sends the number of lives left in the game, which is then used to
+            mark the end of an episode during training.
[…]

### 2017-12-18 games.py

@@ -1,0 +1,28 @@
+
+import gym
+
+class Games:
+
+    def __init__(self):   
+
+        # Cool (easy) Atari environments available
+        self.ASTEROIDS = "Asteroids-v0"
+        self.PONG = "Pong-v0"
+        self.SPACE_INVADERS = "SpaceInvaders-v0"
+        self.TENNIS = "Tennis-v0"
+
+        self.available_games_list = [self.ASTEROIDS, self.PONG, self.SPACE_INVADERS, self.TENNIS]
+        self.action_space = {}
+        self.define_action_spaces()
+
+
+    def define_action_spaces(self):
+        
+        for game in self.available_games_list:
+            dummy_env = gym.envs.make(game)
+            self.action_space[game] = dummy_env.action_space.shape[0]
+
+    def get_action_space(self, game):
+        return(self.action_space[game])
+
+        

### 2017-12-18 history.py

@@ -1,0 +1,43 @@
+
+"""
+[Article] The function w from algorithm ϕ described below applies this preprocess-
+ing to the m most recent frames and stacks them to produce the input to the
+Q-function, in which m = 4, although the algorithm is robust to different values of
+m (for example, 3 or 5).
+"""
+
+import numpy as np
+from parameters import Parameters
+
+class Frames_History:
+
+    """ The frame history will be used as an input to the convolutional neural network """
+
+    def __init__(self):
+            
+        """ /!\ Always in Float32 otherwise cv2.resize shits itself /!\ """
+        self.frames_history = np.zeros((Parameters.m_recent_frames, Parameters.image_height, Parameters.image_width), dtype=np.float32)
+
+
+    def add_frame(self, frame):
+
+        """
+        Add a frame to the frame history
+        Sadly, custom slicing is faster than np.roll()
+        Maybe a faster version can be achieved by using np.lib.stride_tricks.as_strided 
+        """
+        self.frames_history[:-1] = self.frames_history[1:]
+        self.frames_history[-1] = frame
+
+
+    def reset(self):
+        self.frames_history *= 0
+
+    
+    def get(self):
+        return(self.frames_history)
+
+
+
+        
+        

### 2017-12-18 parameters.py

@@ -1,0 +1,52 @@
+
+from games import Games
+
+class Parameters:
+
+    # environment settings
+    games = Games()
+    game = games.SPACE_INVADERS
+    action_space = games.get_action_space(game)
+    display = True
+    FPS = 30
+
+    """
+    [Article] Following previous approaches to playing Atari 2600 games, we also use a simple
+    frame-skipping technique. More precisely, the agent sees and selects actions on
+    every kth frame instead of every frame, and its last action is repeated on skipped
+    frames. Because running the emulator forward for one step requires much less
+    computation than having the agent select an action, this technique allows the agent
+    to play roughly k times more games without significantly increasing the runtime.
+    We use k = 4 for all games
+    """
+    frame_skipping = 4
+
+
+    # preprocessing
+    image_width = 84
+    image_height = 84
+
+
+    # Q-learning settings
+    """
+    [Article] As the scale of scores varies greatly from game to game, we clipped all posi-
+    tive rewards at 1 and all negative rewards at 21, leaving 0 rewards unchanged.
+    Clipping the rewards in this manner limits the scale of the error derivatives and
+    makes it easier to use the same learning rate across multiple games. At the same time,
+    it could affect the performance of our agent since it cannot differentiate between
+    rewards of different magnitude.
+    """
+    positive_reward = 1
+    negative_reward = -1
+    no_reward = 0
+
+    """
+    [Article] The function w from algorithm ϕ described below applies this preprocess-
+    ing to the m most recent frames and stacks them to produce the input to the
+    Q-function, in which m = 4, although the algorithm is robust to different values of
+    m (for example, 3 or 5).
+    """
+    m_recent_frames = 4
+
+    MAX_STEPS = 5000
+

### 2017-12-18 tf_decorator.py

@@ -1,0 +1,39 @@
+
+import functools
+import tensorflow as tf
+
+def doublewrap(function):
+    """
+    A decorator decorator, allowing to use the decorator to be used without
+    parentheses if no argument is provided. All arguments must be optional.
+    """
+    @functools.wraps(function)
+    def decorator(*args, **kwargs):
+        if len(args) == 1 and len(kwargs) == 0 and callable(args[0]):
+            return function(args[0])
+        else:
+            return lambda wrapee: function(wrapee, *args, **kwargs)
+    return decorator
+
+@doublewrap
+def define_scope(function, scope=None, *args, **kwargs):
+    """
+    A decorator for functions that define TensorFlow operations. The wrapped
+    function will only be executed once. Subsequent calls to it will directly
+    return the result so that operations are added to the graph only once.
+    The operations added by the function live within a tf.variable_scope(). If
+    this decorator is used with arguments, they will be forwarded to the
+    variable scope. The scope name defaults to the name of the wrapped
+    function.
+    """
+    attribute = '_cache_' + function.__name__
+    name = scope or function.__name__
+    @property
+    @functools.wraps(function)
+    def decorator(self):
+        if not hasattr(self, attribute):
+            with tf.variable_scope(name, *args, **kwargs):
+                setattr(self, attribute, function(self))
+        return getattr(self, attribute)
+    return decorator
+

### 2017-12-18 utils.py

@@ -2,8 +2,34 @@
 # utils.py: Hyper-parameters and stuff
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
+import numpy as np
+import cv2
+from parameters import Parameters
 
-class Params:
-    FPS = 30
-    MAX_STEPS = 5000
+    
+def y_channel(image):
+    """
+    Y-channel = luminance = greyscale = (0.299*R + 0.587*G + 0.114*B)
+    orb.essex.ac.uk/ce/ce316/opencv-intro.pdf
+    """
+    return(np.dot(image[...,:3], [0.299, 0.587, 0.114]))
 
+
+def screen_resize(image):
+    """
+    Resize the image with the width and height of the parameters (84x84 in the article)
+    Using cv2 and INTER_AREA for speed
+    Original size: 210x160x3
+    """
+    return(cv2.resize(image, (Parameters.image_width, Parameters.image_height), interpolation= cv2.INTER_AREA))
+
+
+def remove_flickering(previous_image, image):
+    """
+    [Article] First, to encode a single frame we take the maximum value for each pixel colour
+    value over the frame being encoded and the previous frame. This was necessary to
+    remove flickering that is present in games where some objects appear only in even
+    frames while other objects appear only in odd frames, an artefact caused by the
+    limited number of sprites Atari 2600 can display at once.
+    """
+    return(np.asarray([previous_image, image]).max(axis=0))

### 2017-12-18 agent.py
Load hyper-parameters from config json file

@@ -2,7 +2,6 @@
 # dqn.py
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-from parameters import Parameters
 from environment import Environment
 from dqn import DQN
 

### 2017-12-18 dqn.py

@@ -25,14 +25,14 @@ class DQN:
         produced by the preprocessing map ϕ
         """
         # reshape input to 4d tensor [batch, height, width, channels]
-        input = tf.reshape(self.image, [-1, Parameters.image_height, Parameters.image_width, Parameters.m_recent_frames])
+        input = tf.reshape(self.image, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES])
 
         # convolutional layer 1
         """
         [Article] The first hidden layer convolves 32 filters of 8 x 8 with stride 4 with the
         input image and applies a rectifier nonlinearity
         """
-        W_conv1 = self.weight_variable([8, 8, Parameters.m_recent_frames, 32])
+        W_conv1 = self.weight_variable([8, 8, Parameters.M_RECENT_FRAMES, 32])
         b_conv1 = self.bias_variable([32])
         conv1 = tf.nn.conv2d(input, W_conv1, strides=[1, 4, 4, 1], padding='VALID') # would 'SAME' also work ?
         h_conv1 = tf.nn.relu(conv1 + b_conv1)
@@ -72,13 +72,13 @@ class DQN:
         h_fc1 = tf.nn.relu(fc1 + b_fc1)
 
         # fully connected layer 2 (output layer)
-        W_fc2 = self.weight_variable([512, Parameters.action_space])
-        b_fc2 = self.bias_variable([Parameters.action_space])
+        W_fc2 = self.weight_variable([512, Parameters.ACTION_SPACE])
+        b_fc2 = self.bias_variable([Parameters.ACTION_SPACE])
         fc2 = tf.matmul(h_fc1, W_fc2)
 
         net_output = fc2 + b_fc2
 
-        # network output is of shape (1, Parameters.action_space)
+        # network output is of shape (1, Parameters.ACTION_SPACE)
         
         return(net_output)
     

### 2017-12-18 environment.py

@@ -13,7 +13,7 @@ class Environment:
 
     def __init__(self):
 
-        self.environment = gym.envs.make(Parameters.game)
+        self.environment = gym.envs.make(Parameters.GAME)
         self.new_game()
 
         # let's have a look at the first cnn input
@@ -44,7 +44,7 @@ class Environment:
         """ initialize the history by pushing the first screen 4 (m_recent_frames) times """
 
         self.history = Frames_History()
-        for _ in range(Parameters.m_recent_frames):
+        for _ in range(Parameters.M_RECENT_FRAMES):
             self.history.add_frame(self.screen)
 
     
@@ -75,17 +75,17 @@ class Environment:
     def act(self, action):
 
         """
-        Take the action a certain number of times (Parameters.frame_skipping)
+        Take the action a certain number of times (Parameters.FRAME_SKIPPING)
         as described in the article
         Return the environment state after taking the action x times
         """
 
         lives_before_action = self.lives
-        cumulated_reward = Parameters.no_reward
+        cumulated_reward = Parameters.NO_REWARD
 
         # frame skipping (see Parameters for more information)
         skipped = 0 # break is evil
-        while(skipped < Parameters.frame_skipping and not self.terminal):
+        while(skipped < Parameters.FRAME_SKIPPING and not self.terminal):
             
             self.take_action(action)
             cumulated_reward += self.reward
@@ -98,7 +98,7 @@ class Environment:
             mark the end of an episode during training.
             """
             if self.lives < lives_before_action:
-                cumulated_reward -= Parameters.negative_reward
+                cumulated_reward -= Parameters.NEGATIVE_REWARD
                 self.terminal = True
             
             skipped += 1
@@ -128,6 +128,6 @@ class Environment:
         """
         Display the game on screen only if the display parameter is True
         """
-        if(Parameters.display):
+        if(Parameters.DISPLAY):
             self.environment.render()
 

### 2017-12-18 history.py

@@ -16,7 +16,7 @@ class Frames_History:
     def __init__(self):
             
         """ /!\ Always in Float32 otherwise cv2.resize shits itself /!\ """
-        self.frames_history = np.zeros((Parameters.m_recent_frames, Parameters.image_height, Parameters.image_width), dtype=np.float32)
+        self.frames_history = np.zeros((Parameters.M_RECENT_FRAMES, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float32)
 
 
     def add_frame(self, frame):

### 2017-12-18 main.py

@@ -3,7 +3,7 @@
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 from dqn import Agent
-from utils import Params
+from parameters import Parameters
 
 import gym
 import time
@@ -11,16 +11,10 @@ import numpy as np
 import tensorflow as tf
 
 
-# Cool (easy) Atari environments
-ASTEROIDS = "Asteroids-v0"
-PONG = "Pong-v0"
-SPACE_INVADERS = "SpaceInvaders-v0"
-TENNIS = "Tennis-v0"
-
-
 if __name__ == "__main__":
+    Parameters.load("parameters.json")
     with tf.Session() as session:
-        environment = gym.envs.make(SPACE_INVADERS)
+        environment = gym.envs.make(Parameters.SPACE_INVADERS)
         actions = environment.action_space
         assert(type(actions) == gym.spaces.discrete.Discrete)
         agent = Agent(session, actions.shape[0])
@@ -28,13 +22,13 @@ if __name__ == "__main__":
         while True:
             s = environment.reset()
             environment.render(mode='human')
-            for t in range(Params.MAX_STEPS):
+            for t in range(Parameters.MAX_STEPS):
                 a = agent.select_action(s)
                 
                 s_prime, r, done, info = environment.step(a)
                 if done:
                     break # End of episode
-                time.sleep(1.0/Params.FPS) # Wait one step
+                time.sleep(1.0 / Parameters.FPS) # Wait one step
                 environment.render()
 
             s = s_prime

### 2017-12-18 parameters.py

@@ -4,49 +4,32 @@ from games import Games
 class Parameters:
 
     # environment settings
-    games = Games()
-    game = games.SPACE_INVADERS
-    action_space = games.get_action_space(game)
-    display = True
-    FPS = 30
+    GAMES = Games()
+    GAME = games.SPACE_INVADERS
+    ACTION_SPACE = games.get_action_space(game)
+    DISPLAY = True
 
-    """
-    [Article] Following previous approaches to playing Atari 2600 games, we also use a simple
-    frame-skipping technique. More precisely, the agent sees and selects actions on
-    every kth frame instead of every frame, and its last action is repeated on skipped
-    frames. Because running the emulator forward for one step requires much less
-    computation than having the agent select an action, this technique allows the agent
-    to play roughly k times more games without significantly increasing the runtime.
-    We use k = 4 for all games
-    """
-    frame_skipping = 4
+    @staticmethod
+    def add_attr(name, value):
+        """ Statically add a parameter as an attribute
 
+        :param name: str
+            Name of the new attribute
+        :param value: object
+            Value of the corresponding hyper-parameter
+        """
+        name = name.upper()
+        setattr(Parameters, name, value)
 
-    # preprocessing
-    image_width = 84
-    image_height = 84
+    @staticmethod
+    def load(filepath):
+        """ Statically loads the hyper-parameters from a json file
 
-
-    # Q-learning settings
-    """
-    [Article] As the scale of scores varies greatly from game to game, we clipped all posi-
-    tive rewards at 1 and all negative rewards at 21, leaving 0 rewards unchanged.
-    Clipping the rewards in this manner limits the scale of the error derivatives and
-    makes it easier to use the same learning rate across multiple games. At the same time,
-    it could affect the performance of our agent since it cannot differentiate between
-    rewards of different magnitude.
-    """
-    positive_reward = 1
-    negative_reward = -1
-    no_reward = 0
-
-    """
-    [Article] The function w from algorithm ϕ described below applies this preprocess-
-    ing to the m most recent frames and stacks them to produce the input to the
-    Q-function, in which m = 4, although the algorithm is robust to different values of
-    m (for example, 3 or 5).
-    """
-    m_recent_frames = 4
-
-    MAX_STEPS = 5000
-
+        :param filepath: str
+            Path to the json parameter file
+        """
+        with open(filepath, "r") as f:
+            data = json.load(f)
+            for key in data.keys():
+                if type(data[key]) == dict:
+                    Parameters.add_attr(key, data[key]["value"])

### 2017-12-18 utils.py

@@ -21,7 +21,7 @@ def screen_resize(image):
     Using cv2 and INTER_AREA for speed
     Original size: 210x160x3
     """
-    return(cv2.resize(image, (Parameters.image_width, Parameters.image_height), interpolation= cv2.INTER_AREA))
+    return(cv2.resize(image, (Parameters.IMAGE_WIDTH, Parameters.IMAGE_HEIGHT), interpolation= cv2.INTER_AREA))
 
 
 def remove_flickering(previous_image, image):

### 2017-12-18 agent.py
Minor changes

@@ -1,5 +1,5 @@
 # -*- coding: utf-8 -*-
-# dqn.py
+# agent.py
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 from environment import Environment
@@ -20,5 +20,3 @@ class Agent:
         # The agent uses its experience and expertise acquired
         # through deep learning to make intelligent actions
         return np.random.randint(0, self.n_actions, size=1)[0]
-
-Agent()

### 2017-12-18 environment.py

@@ -1,10 +1,14 @@
+# -*- coding: utf-8 -*-
+# environment.py 
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-import gym
-import numpy as np
-import cv2
 from history import Frames_History
 from parameters import Parameters
 import utils
+
+import cv2
+import gym
+import numpy as np
 import matplotlib.pyplot as plt
 
 # ϕ
@@ -32,7 +36,7 @@ class Environment:
         self.take_action(first_action)
         self.render()
         self.reward = 0
-        self.terminal = False # True?
+        self.terminal = False # True? -> Parameter ?
 
         self.initialize_screens_history()
 

### 2017-12-18 history.py

@@ -1,3 +1,6 @@
+# -*- coding: utf-8 -*-
+# history.py  : Experience replay
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 """
 [Article] The function w from algorithm ϕ described below applies this preprocess-

### 2017-12-18 parameters.py

@@ -1,3 +1,6 @@
+# -*- coding: utf-8 -*-
+# parameters.py : Load parameters from json config file
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 from games import Games
 
@@ -11,7 +14,9 @@ class Parameters:
 
     @staticmethod
     def add_attr(name, value):
-        """ Statically add a parameter as an attribute
+        """ Statically adds a parameter as an attribute
+        to class Parameters. All new Parameters atributes
+        are in capital letters.
 
         :param name: str
             Name of the new attribute

### 2017-12-18 tf_decorator.py

@@ -1,3 +1,6 @@
+# -*- coding: utf-8 -*-
+# tf_decorator.py 
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 import functools
 import tensorflow as tf

### 2017-12-18 utils.py

@@ -1,5 +1,5 @@
 # -*- coding: utf-8 -*-
-# utils.py: Hyper-parameters and stuff
+# utils.py: Image preprocessing
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 import numpy as np

### 2017-12-23 article/article.tex
[Article] Nothing of much interest

@@
   by learning only from visual features. Despite the current progress, learning to take complex decisions on the basis of high-dimensional visual data remains on
-  ongoing challenge.
+  ongoing challenge. In particular, because of the instability of neural networks, we also explored state-of-the-art dueling architectures to enhance performance
+  by separating the high-level representation of states and action values (rewards).
 
@@
 Since we are handling visual features, we used screen frames as environment state representations. Each frame is a (210 x 160) color image sampled at
-a frequency of 60 Hz.
+a frequency of 60 Hz. We denote the environment state at current step $t$ by $s_t$. The agent gathers experience by incorporating new experience samples
+to its history. An experience is a tuple $e_t = (s_t, a_t, r_t, s_{t+1})$ appended to the experience history immediately after being generated.
+To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse on the experience
+replay algorithm. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
+We compared this technique to a more efficient approach called prioritized experience replay \citep{DBLP:journals/corr/SchaulQAS15}.
 
   \todo{Antoine suite}
+
+It has been shown that DQN has an ability to learn by acquiring experience generated from policies other than its own.
 

### 2017-12-23 article/article.tex
[Article] Add a Makefile
[Article] Type + Add gitignore

@@
 
+\usepackage{hyperref}
+
 \newcommand\todo[1]{\textcolor{red}{TODO: #1}}
+
+\newcommand{\E}{\mathbb E}
 
@@
 \[
-  Q^{*}(s, a) = \max_{\pi} \mathop{\mathbb{E}}[ r_t + \gamma r_{t+1} + \gamma^2 r_{t+2} + ... | s_t = s, a_t = a, \pi]
+  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k | s_t = s, a_t = a, \pi\right].
 \]
 where $\pi = P(a | s)$ is a behavior policy, $t$ is the current step, $r_t$ is the reward for step $t$, $a_t$ is the action selected at time $t$,
-and $\gamma$ is the discount factor. Thus, we can described the Q-value as the present value of all future rewards at current step $t$,
+and $\gamma$ is the discount factor. Thus, we can describe the Q-value as the present value of all future rewards at current step $t$,
 with respect to the environment state and the action selected at $t$.

### 2017-12-24 agent.py
DQN: finished optimization, huber loss, added setter for target dqn updates and getter in all layers and parameters | replay memory done | Agent: added DQN update method and batch q learning | made only one class for DQN and Target DQN unlike the other github implementation

@@ -3,20 +3,118 @@
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 from environment import Environment
+from memory import Memory
+from parameters import Parameters
 from dqn import DQN
 
+import random
 import numpy as np
 import tensorflow as tf
 
 
 class Agent:
-    def __init__(self):
+
+    def __init__(self, environment):
         
-        self.environment = Environment()
-        self.dqn = DQN(self.environment.get_input())
+        self.environment = environment
+        self.memory = Memory()
+        self.step = 0
+        
+        # initialize the DQN and target DQN (with respective placeholders)
+        self.dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES], name = "DQN_input")
+        self.dqn = DQN(self.dqn_input)
+
+        self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES], name = "target_DQN_input")
+        self.target_dqn = DQN(self.target_dqn_input)
+
+        # initialize the tensorflow session and variables
+        self.tf_session = tf.Session()
+        self.tf_session.run(tf.global_variables_initializer())
 
 
-    def select_action(self, s):
-        # The agent uses its experience and expertise acquired
-        # through deep learning to make intelligent actions
-        return np.random.randint(0, self.n_actions, size=1)[0]
+    
+    def batch_q_learning(self):
+        """
+        Apply Q-learning updates, or minibatch updates, to samples of experience,
+        (s, a, r, s') ~ U(D), drawn at random from the pool of stored samples.
+        """
+
+        if(self.memory.get_usage() > Parameters.agent_history_length):
+            
+            state_t, action, reward, state_t_plus_1, terminal = self.memory.bring_back_memories()
+
+            q_t_plus_1 = self.tf_session.run(self.target_dqn.q_values, {self.target_dqn_input: state_t_plus_1})
+            max_q_t_plus_1 = np.max(q_t_plus_1, axis=1)
+
+            target_q_t = (1. - terminal) * Parameters.discount_factor * max_q_t_plus_1 + reward
+
+            _, q_t, loss = self.tf_session.run([self.dqn.optimize, self.dqn.q_values, self.dqn.error],
+            {
[…]

### 2017-12-24 dqn.py

@@ -6,26 +6,39 @@ from parameters import Parameters
 
 class DQN:
 
-    def __init__(self, image):
+    def __init__(self, state):
 
-        # receiving image placeholder
-        self.image = image
+        # receiving state placeholder
+        self.state = state
+
+        # weights and biases dictionary
+        self.learning_parameters = {}
+        self.layers : {}
 
         # initialize tensorflow graph
-        self.prediction
-        self.train
+        self.q_values
+        self.optimize
         self.error
+
+        # Initialize input placeholder to assign values to weights and biases
+        # Used when we transfert them from the training network to the target network
+        # or when we load DQN parameters previously saved in a file
+        self.l_param_input = {}
+        self.assign_operator = {}
+        for variable_name in self.learning_parameters.keys():
+            self.l_param_input[variable_name] = tf.placeholder(tf.float32, self.learning_parameters[variable_name].get_shape().as_list(), name=variable_name)
+            self.assign_operator[variable_name] = self.learning_parameters[variable_name].assign(self.l_param_input[variable_name])
 
 
     @define_scope
-    def prediction(self):
+    def q_values(self):
         
         """
-        [Article] The input to the neural network consists of an 84 x 84 x 4 image 
+        [Article] The input to the neural network (the state) consists of an 84 x 84 x 4 image 
         produced by the preprocessing map ϕ
         """
         # reshape input to 4d tensor [batch, height, width, channels]
-        input = tf.reshape(self.image, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES])
+        input = tf.reshape(self.state, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES])
 
         # convolutional layer 1
         """
@@ -62,7 +75,7 @@ class DQN:
         h_conv3 = tf.nn.relu(conv3 + b_conv3)
 
         # output of conv 3 is of shape [-1 x 7 x 7 x 64]
-       
+        
         h_conv3_flat = tf.reshape(h_conv3, [-1, 7 * 7 * 64])
 
         # fully connected layer 1
@@ -76,8 +89,6 @@ class DQN:
         b_fc2 = self.bias_variable([Parameters.ACTION_SPACE])
         fc2 = tf.matmul(h_fc1, W_fc2)
 
-        net_output = fc2 + b_fc2
-
         # network output is of shape (1, Parameters.ACTION_SPACE)
         """
[…]

### 2017-12-24 history.py

@@ -19,7 +19,7 @@ class Frames_History:
     def __init__(self):
             
         """ /!\ Always in Float32 otherwise cv2.resize shits itself /!\ """
-        self.frames_history = np.zeros((Parameters.M_RECENT_FRAMES, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float32)
+        self.frames_history = np.zeros((Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float32)
 
 
     def add_frame(self, frame):

### 2017-12-24 memory.py

@@ -1,0 +1,96 @@
+
+from parameters import Parameters
+import numpy as np
+
+class Memory:
+
+    def __init__(self):
+
+        self.memory_size = Parameters.replay_memory_size
+        self.current_memory_index = 0
+        self.memory_usage = 0
+
+        self.actions = np.empty(self.memory_size, dtype=np.uint8)
+        self.rewards = np.empty(self.memory_size, dtype=np.integer)
+        self.screens = np.empty(self.memory_size, Parameters.image_height, Parameters.image_width, dtype=np.float16)
+        self.terminals = np.empty(self.memory_size, dtype=np.bool)
+
+        self.minibatch_size = Parameters.minibatch_size
+
+        self.state_t = np.empty((self.minibatch_size, Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float16)
+        self.state_t_plus_1 = np.empty((self.minibatch_size, Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float16)
+
+
+    def add(self, screen, action, reward, terminal):
+
+        self.actions[self.current_memory_index] = action
+        self.rewards[self.current_memory_index] = reward
+        self.screens[self.current_memory_index, ...] = screen
+        self.terminals[self.current_memory_index] = terminal
+
+        self.memory_usage = max(self.memory_usage, self.current_memory_index + 1)
+        self.current_memory_index = (self.current_memory_index + 1) % self.memory_size
+
+    
+    def get_state(self, state_index):
+
+        state = None
+
+        if(not self.memory_usage):
+            print("Memory is empty")
+        else:
+
+            state_index = state_index % self.memory_usage
+
+            if(state_index >= Parameters.agent_history_length - 1):
+                state = self.screens[(state_index - Parameters.agent_history_length) + 1 : state_index + 1, ...]
+            else:
+                # negative indices don't work well with slices in numpy..
+                state = self.screens[np.array([(state_index-i) % self.memory_usage for i in reversed(range(Parameters.agent_history_length))]), ...]
+
+        return(state)
+        
+    
+    def bring_back_memories(self):
+        """
+        [Article] our algorithm only stores the last N experience tuples in the replay
+        memory, and samples uniformly at random from D when performing updates. This
+        approach is in some respects limited because the memory buffer does not differ-
+        entiate important transitions and always overwrites with recent transitions owing
+        to the finite memory size N. Similarly, the uniform sampling gives equal impor-
[…]

### 2017-12-24 agent.py
Uppercase parameters

@@ -39,14 +39,14 @@ class Agent:
         (s, a, r, s') ~ U(D), drawn at random from the pool of stored samples.
         """
 
-        if(self.memory.get_usage() > Parameters.agent_history_length):
+        if(self.memory.get_usage() > Parameters.AGENT_HISTORY_LENGTH):
             
             state_t, action, reward, state_t_plus_1, terminal = self.memory.bring_back_memories()
 
             q_t_plus_1 = self.tf_session.run(self.target_dqn.q_values, {self.target_dqn_input: state_t_plus_1})
             max_q_t_plus_1 = np.max(q_t_plus_1, axis=1)
 
-            target_q_t = (1. - terminal) * Parameters.discount_factor * max_q_t_plus_1 + reward
+            target_q_t = (1. - terminal) * Parameters.DISCOUNT_FACTOR * max_q_t_plus_1 + reward
 
             _, q_t, loss = self.tf_session.run([self.dqn.optimize, self.dqn.q_values, self.dqn.error],
             {
@@ -64,14 +64,14 @@ class Agent:
         self.memory.add(screen, action, reward, terminal)
 
         # if we started learning
-        if(self.step > Parameters.replay_start_size):
+        if(self.step > Parameters.REPLAY_START_SIZE):
 
             # If we are not between successive SGD updates
-            if(not(self.step % Parameters.update_frequency)):
+            if(not(self.step % Parameters.UPDATE_FREQUENCY)):
                 self.batch_q_learning()
             
             # If we are not between successive target network update
-            if(not(self.step % Parameters.target_network_update_frequency)):
+            if(not(self.step % Parameters.TARGET_NETWORK_UPDATE_FREQUENCY)):
                 self.update_target_dqn()
             
 
@@ -83,7 +83,9 @@ class Agent:
         """
 
         # compute epsilon at step t
-        eps = Parameters.final_exploration + max(0., (Parameters.initial_exploration - Parameters.final_exploration) * (Parameters.final_exploration_frame - max(0., self.step - Parameters.replay_start_size)) / Parameters.final_exploration_frame)
+        dt_final = Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION
+        dt = self.step - Parameters.REPLAY_START_SIZE
+        eps = Parameters.FINAL_EXPLORATION + max(0., dt * (Parameters.FINAL_EXPLORATION_FRAME - max(0., dt)) / Parameters.FINAL_EXPLORATION_FRAME)
 
         if random.random() < eps:
             # take a random action

### 2017-12-24 dqn.py

@@ -129,9 +129,9 @@ class DQN:
     @define_scope
     def optimize(self):
         
-        self.optimizer = tf.train.RMSPropOptimizer( learning_rate = Parameters.learning_rate, 
-                                                    momentum = Parameters.gradient_momentum,
-                                                    epsilon = Parameters.min_squared_gradient )
+        self.optimizer = tf.train.RMSPropOptimizer( learning_rate = Parameters.LEARNING_RATE, 
+                                                    momentum = Parameters.GRADIENT_MOMENTUM,
+                                                    epsilon = Parameters.MIN_SQUARED_GRADIENT )
         
         return(optimizer.minimize(self.error))
 

### 2017-12-24 history.py

@@ -19,7 +19,7 @@ class Frames_History:
     def __init__(self):
             
         """ /!\ Always in Float32 otherwise cv2.resize shits itself /!\ """
-        self.frames_history = np.zeros((Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float32)
+        self.frames_history = np.zeros((Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float32)
 
 
     def add_frame(self, frame):

### 2017-12-24 memory.py

@@ -6,19 +6,19 @@ class Memory:
 
     def __init__(self):
 
-        self.memory_size = Parameters.replay_memory_size
+        self.memory_size = Parameters.REPLAY_MEMORY_SIZE
         self.current_memory_index = 0
         self.memory_usage = 0
 
         self.actions = np.empty(self.memory_size, dtype=np.uint8)
         self.rewards = np.empty(self.memory_size, dtype=np.integer)
-        self.screens = np.empty(self.memory_size, Parameters.image_height, Parameters.image_width, dtype=np.float16)
+        self.screens = np.empty(self.memory_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, dtype=np.float16)
         self.terminals = np.empty(self.memory_size, dtype=np.bool)
 
-        self.minibatch_size = Parameters.minibatch_size
+        self.minibatch_size = Parameters.MINIBATCH_SIZE
 
-        self.state_t = np.empty((self.minibatch_size, Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float16)
-        self.state_t_plus_1 = np.empty((self.minibatch_size, Parameters.agent_history_length, Parameters.image_height, Parameters.image_width), dtype=np.float16)
+        self.state_t = np.empty((self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
+        self.state_t_plus_1 = np.empty((self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
 
 
     def add(self, screen, action, reward, terminal):
@@ -42,11 +42,11 @@ class Memory:
 
             state_index = state_index % self.memory_usage
 
-            if(state_index >= Parameters.agent_history_length - 1):
-                state = self.screens[(state_index - Parameters.agent_history_length) + 1 : state_index + 1, ...]
+            if(state_index >= Parameters.AGENT_HISTORY_LENGTH - 1):
+                state = self.screens[(state_index - Parameters.AGENT_HISTORY_LENGTH) + 1 : state_index + 1, ...]
             else:
                 # negative indices don't work well with slices in numpy..
-                state = self.screens[np.array([(state_index-i) % self.memory_usage for i in reversed(range(Parameters.agent_history_length))]), ...]
[…]

### 2017-12-24 article/article.tex
[Article] Action space, normalized performance, preprocessing

@@
 to its history. An experience is a tuple $e_t = (s_t, a_t, r_t, s_{t+1})$ appended to the experience history immediately after being generated.
-To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse on the experience
+To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse to the experience
 replay algorithm. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
 We compared this technique to a more efficient approach called prioritized experience replay \citep{DBLP:journals/corr/SchaulQAS15}.
+To emphasize the importance of the memory replay mechanism, we compared the results obtained using prioritized experience replay, regular
+experience replay, and no experience replay.
 
-  \todo{Antoine suite}
+Every 4 steps of the Q-learning, the agent builds a minibatch from its memory (either one of the three mechanisms we just mentioned):
+this minibatch is used by the DQN to process one step of stochastic gradient descent. In this way the values of Q are estimated iteratively and
+refined over time. The deep neural network incorporates minimal prior knowledge about the environment: the only assumption lies in the number of 
+outputs in the last hidden layer, which has been set as the number of actions. In every Atari game, the action space is discrete and finite: this allows
+us to determine the output shape of the neural network depending on the game itself. The number of actions made available by the environment
+cannot be inferred by the DQN during training phase.
 
-It has been shown that DQN has an ability to learn by acquiring experience generated from policies other than its own.
+Also we compared DQN, as well as other state-of-the-art deep reinforcement learning architectures, to simpler reinforcement learning techniques (including
+random action selection) using a normalized performance metric. The latter is formalized as below:
+\[
+    P = 100 \times (score - RAS) / (HS - RAS)
+\]
+where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
+Space Invaders professional gamer \textbf{Stanislas Gueniffey}. By normalizing in this manner, the random action selection score RAS results in
+a normalized performance of 0\% and the human score results in a normalized performance of 100\%.
+
+Finally it has been pointed out by the original DQN conceptors that DQN has an ability to learn by acquiring experience generated from policies
+other than its own. \todo{On peut en faire qqch pas vrai ?}
+
+\todo{introduce DDDQN}
 
@@
 
-  \todo{pour Robin}
+Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}
 
-  \todo{Citer ALE} \citep{bellemare13arcade}
+We developed our models with the support of the Arcade Learning Environment \citep{bellemare13arcade}. This gives the opportunity to either
+retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
+applications). Because of the very high dimensionality of raw images, we applied some preprocessing steps on them. These steps are also aimed at 
+removing some artifacts produced by the Atari architecture. More specifically, the flickering effect implies that some game sprites only appear
+in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
+Then the sample size has been reduced by a factor of three by mapping RBG values to the pixel luminance, using the following formula:
+\[
+    luminance = 0.299 \times R + 0.587 \times G + 0.114 \times B
+\]
+After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
+$84 \times 84$ pixel grid.
+
+\todo{Suite: CNN architecture}
 

### 2017-12-24 agent.py
finish observe method + minor changes, Merry Christmasssssssssssssgit pull!

@@ -2,6 +2,7 @@
 # agent.py
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
+from utils import reward_clipper
 from environment import Environment
 from memory import Memory
 from parameters import Parameters
@@ -57,20 +58,27 @@ class Agent:
 
 
     def observe(self, screen, action, reward, terminal):
+        """
+        [Article] The agent observes an image from the emulator, 
+        which is a vector of pixel values representing the current screen. 
+        In addition it receives a reward representing the change in game score. 
+
+        Updates the environment's history and agent's memory and performs an SGD update
+        and/or updates the Target DQN 
+        """
         
         # update agent's memory and environment's history
-
-        # add screen to history (TODO double check the best way to do it)
-        self.memory.add(screen, action, reward, terminal)
+        self.environment.add_current_screen_to_history()
+        self.memory.add(screen, action, reward_clipper(reward), terminal)
 
         # if we started learning
         if(self.step > Parameters.REPLAY_START_SIZE):
 
-            # If we are not between successive SGD updates
+            # Perform SGD updates at frequency [Parameters.UPDATE_FREQUENCY]
             if(not(self.step % Parameters.UPDATE_FREQUENCY)):
                 self.batch_q_learning()
             
-            # If we are not between successive target network update
+            # Update Target DQN at frequency [Parameters.TARGET_NETWORK_UPDATE_FREQUENCY]
             if(not(self.step % Parameters.TARGET_NETWORK_UPDATE_FREQUENCY)):
                 self.update_target_dqn()
             

### 2017-12-24 environment.py

@@ -94,8 +94,6 @@ class Environment:
             self.take_action(action)
             cumulated_reward += self.reward
 
-            ## TODO ?? Should we update screen history or not here ?? ##
-
             """
             [Article] For games where there is a life counter, the Atari
             2600 emulator also sends the number of lives left in the game, which is then used to

### 2017-12-24 utils.py

@@ -33,3 +33,15 @@ def remove_flickering(previous_image, image):
     limited number of sprites Atari 2600 can display at once.
     """
     return(np.asarray([previous_image, image]).max(axis=0))
+
+
+def reward_clipper(reward):
+    """
+    [Article] As the scale of scores varies greatly from game to game, we clipped 
+    all positive rewards at 1 and all negative rewards at 21, leaving 0 rewards unchanged. 
+    Clipping the rewards in this manner limits the scale of the error derivatives and 
+    makes it easier to use the same learning rate across multiple games. 
+    At the same time, it could affect the performance of our agent since it cannot 
+    differentiate between rewards of different magnitude.
+    """
+    return(max(Parameters.NEGATIVE_REWARD, min(Parameters.POSITIVE_REWARD, reward)))

### 2017-12-25 agent.py
Christmas debugging

@@ -51,9 +51,9 @@ class Agent:
 
             _, q_t, loss = self.tf_session.run([self.dqn.optimize, self.dqn.q_values, self.dqn.error],
             {
-                self.dqn.target_q = target_q_t
-                self.dqn.action = action
-                self.dqn_input = state_t
+                self.dqn.target_q : target_q_t,
+                self.dqn.action : action,
+                self.dqn_input : state_t
             })
 
 

### 2017-12-25 dqn.py

@@ -1,7 +1,12 @@
+# -*- coding: utf-8 -*-
+# dqn.py : Deep Q-Networks
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-import tensorflow as tf
 from tf_decorator import *
 from parameters import Parameters
+
+import tensorflow as tf
+from tensorflow.python.ops import array_ops as tf_array_ops
 
 
 class DQN:
@@ -13,7 +18,7 @@ class DQN:
 
         # weights and biases dictionary
         self.learning_parameters = {}
-        self.layers : {}
+        self.layers = {}
 
         # initialize tensorflow graph
         self.q_values
@@ -35,7 +40,7 @@ class DQN:
         
         """
         [Article] The input to the neural network (the state) consists of an 84 x 84 x 4 image 
-        produced by the preprocessing map ϕ
+        produced by the preprocessing map "phi"
         """
         # reshape input to 4d tensor [batch, height, width, channels]
         input = tf.reshape(self.state, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES])
@@ -133,7 +138,7 @@ class DQN:
                                                     momentum = Parameters.GRADIENT_MOMENTUM,
                                                     epsilon = Parameters.MIN_SQUARED_GRADIENT )
         
-        return(optimizer.minimize(self.error))
+        return(self.optimizer.minimize(self.error))
 
 
     @define_scope
@@ -149,7 +154,7 @@ class DQN:
         # convert the action to one-hot representation in order to compute the error
         action_one_hot = tf.one_hot(self.action, Parameters.ACTION_SPACE, on_value=1, off_value=0, name="action_one_hot")
         
-        self.q_acted = tf.reduce_sum(self.q_values * action_one_hot, axis = 1, name="q_acted")
+        self.q_acted = tf.reduce_sum(self.q_values * tf.cast(action_one_hot, tf.float32), axis = 1, name="q_acted")
         
         self.delta = self.target_q - self.q_acted
 
@@ -166,7 +171,7 @@ class DQN:
         x =  0.5 * x^2                  if |x| <= d
         x =  0.5 * d^2 + d * (|x| - d)  if |x| > d
         """
-        self.clipped_error = tf.array_ops.where(tf.abs(self.delta) < 1.0, 
+        self.clipped_error = tf_array_ops.where(tf.abs(self.delta) < 1.0, 
[…]

### 2017-12-25 environment.py

@@ -6,7 +6,6 @@ from history import Frames_History
 from parameters import Parameters
 import utils
 
-import cv2
 import gym
 import numpy as np
 import matplotlib.pyplot as plt

### 2017-12-25 main.py

@@ -2,7 +2,7 @@
 # main.py 
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-from dqn import Agent
+from agent import Agent
 from parameters import Parameters
 
 import gym
@@ -14,10 +14,10 @@ import tensorflow as tf
 if __name__ == "__main__":
     Parameters.load("parameters.json")
     with tf.Session() as session:
-        environment = gym.envs.make(Parameters.SPACE_INVADERS)
+        environment = gym.envs.make(Parameters.GAME)
         actions = environment.action_space
         assert(type(actions) == gym.spaces.discrete.Discrete)
-        agent = Agent(session, actions.shape[0])
+        agent = Agent(environment)
 
         while True:
             s = environment.reset()

### 2017-12-25 memory.py

@@ -12,13 +12,14 @@ class Memory:
 
         self.actions = np.empty(self.memory_size, dtype=np.uint8)
         self.rewards = np.empty(self.memory_size, dtype=np.integer)
-        self.screens = np.empty(self.memory_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, dtype=np.float16)
+        self.screens = np.empty((self.memory_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
         self.terminals = np.empty(self.memory_size, dtype=np.bool)
 
         self.minibatch_size = Parameters.MINIBATCH_SIZE
 
-        self.state_t = np.empty((self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
-        self.state_t_plus_1 = np.empty((self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
+        state_shape = (self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH)
+        self.state_t = np.empty(state_shape, dtype=np.float16)
+        self.state_t_plus_1 = np.empty(state_shape, dtype=np.float16)
 
 
     def add(self, screen, action, reward, terminal):

### 2017-12-25 parameters.py

@@ -4,12 +4,15 @@
 
 from games import Games
 
+import json
+
+
 class Parameters:
 
     # environment settings
     GAMES = Games()
-    GAME = games.SPACE_INVADERS
-    ACTION_SPACE = games.get_action_space(game)
+    GAME = GAMES.SPACE_INVADERS
+    ACTION_SPACE = GAMES.get_action_space(GAME)
     DISPLAY = True
 
     @staticmethod

### 2017-12-25 utils.py

@@ -2,9 +2,10 @@
 # utils.py: Image preprocessing
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-import numpy as np
-import cv2
 from parameters import Parameters
+
+import numpy as np
+# import cv2
 
     
 def y_channel(image):

### 2017-12-26 dqn.py
quick fixes

@@ -131,7 +131,7 @@ class DQN:
     @define_scope
     def smartest_action(self):
 
-        highest_q_value = tf.argmax(self.q_values, dimension = 1)
+        highest_q_value = tf.argmax(self.q_values, axis=1)
         return(highest_q_value)
 
 
@@ -158,7 +158,7 @@ class DQN:
         # convert the action to one-hot representation in order to compute the error
         action_one_hot = tf.one_hot(self.action, Parameters.ACTION_SPACE, on_value=1, off_value=0, name="action_one_hot")
         
-        self.q_acted = tf.reduce_sum(self.q_values * tf.cast(action_one_hot, tf.float32), axis = 1, name="q_acted")
+        self.q_acted = tf.reduce_sum(self.q_values * tf.cast(action_one_hot, tf.float32), axis=1, name="q_acted")
         
         self.delta = self.target_q - self.q_acted
 

### 2017-12-26 environment.py

@@ -21,9 +21,9 @@ class Environment:
     def __init__(self):
 
         self.environment = gym.envs.make(Parameters.GAME)
-        self.environment = self.environment.unwrapped
-        assert(isinstance(self.environment, AtariEnv))
-        assert(isinstance(self.environment.ale, ALEInterface))
+        self.lives = np.inf
+        # assert(isinstance(self.environment, AtariEnv))
+        # assert(isinstance(self.environment.ale, ALEInterface))
         self.new_game()
 
         # let's have a look at the first cnn input
@@ -80,7 +80,9 @@ class Environment:
 
     def take_action(self, action):
         self._previous_screen = self._screen
-        self._screen, self.reward, self.terminal, _ = self.environment.step(action)
+        self._screen, self.reward, self.terminal, info = self.environment.step(action)
+        self.lives = info["ale.lives"]
+
 
     
     def process_step(self, action):
@@ -90,7 +92,7 @@ class Environment:
         as described in the article
         Return the environment state after taking the action x times
         """
-
+        
         lives_before_action = self.lives
         cumulated_reward = Parameters.NO_REWARD
 
@@ -113,7 +115,6 @@ class Environment:
             skipped += 1
 
         self.reward = cumulated_reward
-        self.render()
 
         return(self.state, self.reward, self.terminal)
     
@@ -121,11 +122,6 @@ class Environment:
     @property
     def screen(self):
         return(utils.preprocess_img(self._previous_screen, self._screen))
-
-
-    @property
-    def lives(self):
-        return(self.environment.ale.lives())
 
     
     @property
@@ -140,3 +136,12 @@ class Environment:
         if(Parameters.DISPLAY):
             self.environment.render(mode=mode)
 
+
+    def reset(self):
+        """
+        Reset the environment only if the number of lives is
+        """
+        if(not self.lives):
+            self.environment.reset()
+
+            

### 2017-12-26 main.py

@@ -14,19 +14,21 @@ import tensorflow as tf
 
 if __name__ == "__main__":
     Parameters.load("parameters.json")
-    with tf.Session() as session:
-        environment = Environment()
-        agent = Agent(environment)
+    environment = Environment()
+    agent = Agent(environment)
 
-        while True:
-            environment.render(mode='human')
-            for t in range(Parameters.MAX_STEPS):
-                a = agent.select_action()
-                
-                s_prime, r, done = environment.process_step(a)
-                if done:
-                    break # End of episode
-                time.sleep(1.0 / Parameters.FPS) # Wait one step
-                environment.render()
+    while True:
+        environment.reset()
+        environment.render(mode='human')
+        environment.terminal = False
+        for t in range(Parameters.MAX_STEPS):
+            a = agent.select_action()
+            
+            s_prime, r, done = environment.process_step(a)
+            
+            if done:
+                break # End of episode
+            time.sleep(1.0 / Parameters.FPS) # Wait one step
+            environment.render()
 
-            s = s_prime
+        s = s_prime

### 2017-12-27 agent.py
functional training method + fix getter/setter of TF layers and parameters + main + major fix in biases (was constant and not casted to Variable [oui j'ai mucho honte])

@@ -9,6 +9,8 @@ from parameters import Parameters
 from dqn import DQN
 
 import random
+import time
+
 import numpy as np
 import tensorflow as tf
 
@@ -22,16 +24,48 @@ class Agent:
         self.step = 0
         
         # initialize the DQN and target DQN (with respective placeholders)
-        self.dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES], name = "DQN_input")
+        self.dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "DQN_input")
         self.dqn = DQN(self.dqn_input)
 
-        self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES], name = "target_DQN_input")
+        self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "target_DQN_input")
         self.target_dqn = DQN(self.target_dqn_input)
 
         # initialize the tensorflow session and variables
         self.tf_session = tf.Session()
         self.tf_session.run(tf.global_variables_initializer())
 
+
+    def train(self):
+
+        self.step = 0
+
+        while(self.step < Parameters.MAX_STEPS):
+
+            self.environment.render(mode='human')
+            self.environment.terminal = False
+            for t in range(Parameters.MAX_STEPS):
+                
+                # select an action
+                action = self.select_action()
+                
+                # process step
+                state, reward, terminal = self.environment.process_step(action)
+                
+                # observe the consequence of the action
+                self.observe(state, action, reward, terminal)
+
+                if terminal:
+                    self.environment.reset()
+                    break
+
+                if(Parameters.SLEEP_BETWEEN_STEPS):
+                    time.sleep(1.0 / Parameters.FPS) # Wait one step
+
+                self.environment.render()
+                
+                self.step += 1
+
+            
 
     
     def batch_q_learning(self):
[…]

### 2017-12-27 dqn.py

@@ -28,15 +28,17 @@ class DQN:
         # Initialize input placeholder to assign values to weights and biases
         # Used when we transfert them from the training network to the target network
         # or when we load DQN parameters previously saved in a file
-        self.l_param_input = {}
-        self.assign_operator = {}
-        for variable_name in self.learning_parameters.keys():
-            self.l_param_input[variable_name] = tf.placeholder(tf.float32, self.learning_parameters[variable_name].get_shape().as_list(), name=variable_name)
+        with tf.variable_scope("test"):
+                
+            self.l_param_input = {}
+            self.assign_operator = {}
+            for variable_name in self.learning_parameters.keys():
+                self.l_param_input[variable_name] = tf.placeholder(tf.float32, self.learning_parameters[variable_name].get_shape().as_list(), name=variable_name)
 
-            try: # If mutable tensor (Variable)
-                self.assign_operator[variable_name] = self.learning_parameters[variable_name].assign(self.l_param_input[variable_name])
-            except AttributeError:
-                pass # TODO: what should we do?
+                try: # If mutable tensor (Variable)
+                    self.assign_operator[variable_name] = self.learning_parameters[variable_name].assign(self.l_param_input[variable_name])
+                except AttributeError as e:
+                    print(e)
 
 
     @define_scope
@@ -47,14 +49,14 @@ class DQN:
         produced by the preprocessing map "phi"
         """
         # reshape input to 4d tensor [batch, height, width, channels]
-        input = tf.reshape(self.state, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.M_RECENT_FRAMES])
+        input = tf.reshape(self.state, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH])
 
         # convolutional layer 1
         """
         [Article] The first hidden layer convolves 32 filters of 8 x 8 with stride 4 with the
         input image and applies a rectifier nonlinearity
         """
-        W_conv1 = self.weight_variable([8, 8, Parameters.M_RECENT_FRAMES, 32])
+        W_conv1 = self.weight_variable([8, 8, Parameters.AGENT_HISTORY_LENGTH, 32])
         b_conv1 = self.bias_variable([32])
[…]

### 2017-12-27 environment.py

@@ -2,17 +2,16 @@
 # environment.py 
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
-from history import FrameHistory
+from history import FramesHistory
 from parameters import Parameters
 import utils
 
 import gym
 from gym.envs.atari.atari_env import AtariEnv
-from atari_py import ALEInterface
+
 import numpy as np
 import matplotlib.pyplot as plt
 
-# ϕ
 
 class Environment:
 
@@ -22,8 +21,6 @@ class Environment:
 
         self.environment = gym.envs.make(Parameters.GAME)
         self.lives = np.inf
-        # assert(isinstance(self.environment, AtariEnv))
-        # assert(isinstance(self.environment.ale, ALEInterface))
         self.new_game()
 
         # let's have a look at the first cnn input
@@ -52,10 +49,10 @@ class Environment:
 
     def initialize_screens_history(self):
 
-        """ initialize the history by pushing the first screen 4 (m_recent_frames) times """
+        """ initialize the history by pushing the first screen 4 (AGENT_HISTORY_LENGTH) times """
 
-        self.history = FrameHistory()
-        for _ in range(Parameters.M_RECENT_FRAMES):
+        self.history = FramesHistory()
+        for _ in range(Parameters.AGENT_HISTORY_LENGTH):
             self.history.add_frame(self.screen)
 
     

### 2017-12-27 history.py

@@ -14,7 +14,7 @@ from parameters import Parameters
 import numpy as np
 
 
-class FrameHistory:
+class FramesHistory:
 
     """ The frame history will be used as an input to the convolutional neural network """
 

### 2017-12-27 main.py

@@ -6,29 +6,13 @@ from agent import Agent
 from environment import Environment
 from parameters import Parameters
 
-import gym
-import time
-import numpy as np
-import tensorflow as tf
-
 
 if __name__ == "__main__":
+    
     Parameters.load("parameters.json")
     environment = Environment()
     agent = Agent(environment)
+    
+    agent.train()
 
-    while True:
-        environment.reset()
-        environment.render(mode='human')
-        environment.terminal = False
-        for t in range(Parameters.MAX_STEPS):
-            a = agent.select_action()
-            
-            s_prime, r, done = environment.process_step(a)
-            
-            if done:
-                break # End of episode
-            time.sleep(1.0 / Parameters.FPS) # Wait one step
-            environment.render()
-
-        s = s_prime
+    

### 2017-12-27 memory.py

@@ -17,7 +17,7 @@ class Memory:
 
         self.minibatch_size = Parameters.MINIBATCH_SIZE
 
-        state_shape = (self.minibatch_size, Parameters.AGENT_HISTORY_LENGTH, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH)
+        state_shape = (self.minibatch_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)
         self.state_t = np.empty(state_shape, dtype=np.float16)
         self.state_t_plus_1 = np.empty(state_shape, dtype=np.float16)
 
@@ -48,6 +48,9 @@ class Memory:
             else:
                 # negative indices don't work well with slices in numpy..
                 state = self.screens[np.array([(state_index-i) % self.memory_usage for i in reversed(range(Parameters.AGENT_HISTORY_LENGTH))]), ...]
+                
+        state = np.swapaxes(state, 0, 1)
+        state = np.swapaxes(state, 1, 2)
 
         return(state)
         

### 2017-12-27 parameters.py

@@ -14,6 +14,7 @@ class Parameters:
     GAME = GAMES.SPACE_INVADERS
     ACTION_SPACE = GAMES.get_action_space(GAME)
     DISPLAY = True
+    SLEEP_BETWEEN_STEPS = True
 
     @staticmethod
     def add_attr(name, value):

### 2017-12-27 agent.py
Define save and restore functions for agent.tf_session

@@ -7,6 +7,7 @@ from environment import Environment
 from memory import Memory
 from parameters import Parameters
 from dqn import DQN
+from os import path, makedirs
 
 import random
 import time
@@ -32,7 +33,26 @@ class Agent:
 
         # initialize the tensorflow session and variables
         self.tf_session = tf.Session()
-        self.tf_session.run(tf.global_variables_initializer())
+        load_session()
+        
+    
+    def load_session(self):
+        save_file = path.join(Parameters.SESSION_SAVE_DIRECTORY, Parameters.SESSION_SAVE_FILENAME)
+        if path.exists(save_file) and path.isfile(save_file):
+                # restore from a previously saved session
+                tf_saver = tf.train.Saver()
+                tf_saver.restore(self.tf_session, save_file)
+        else:
+                # initialize from scratch
+                self.tf_session.run(tf.global_variables_initializer())
+                
+    
+    def save_session(self):
+        save_file = path.join(Parameters.SESSION_SAVE_DIRECTORY, Parameters.SESSION_SAVE_FILENAME)
+        if not path.exists(save_file):
+                makedirs(Parameters.SESSION_SAVE_DIRECTORY)
+        tf_saver = tf.train.Saver()
+        tf_saver.save(self.tf_session, save_file)
 
 
     def train(self):

### 2017-12-27 agent.py
Add save folder and files to parameters + fix save/restore

@@ -33,7 +33,7 @@ class Agent:
 
         # initialize the tensorflow session and variables
         self.tf_session = tf.Session()
-        load_session()
+        self.load_session()
         
     
     def load_session(self):
@@ -49,7 +49,7 @@ class Agent:
     
     def save_session(self):
         save_file = path.join(Parameters.SESSION_SAVE_DIRECTORY, Parameters.SESSION_SAVE_FILENAME)
-        if not path.exists(save_file):
+        if not path.exists(Parameters.SESSION_SAVE_DIRECTORY):
                 makedirs(Parameters.SESSION_SAVE_DIRECTORY)
         tf_saver = tf.train.Saver()
         tf_saver.save(self.tf_session, save_file)
@@ -84,7 +84,11 @@ class Agent:
                 self.environment.render()
                 
                 self.step += 1
-
+                
+                if self.step % 10000 == 0:
+                        self.save_session()
+        
+        self.save_session()
             
 
     

### 2017-12-27 parameters.py

@@ -41,4 +41,7 @@ class Parameters:
             data = json.load(f)
             for key in data.keys():
                 if type(data[key]) == dict:
+                    if key == "SESSION_SAVE_FILENAME":
+                        # make session filename specific to the game being played
+                        data[key]["value"] = data[key]["value"].replace(".", "_"+Parameters.GAME+".")
                     Parameters.add_attr(key, data[key]["value"])

### 2017-12-27 article/article.tex
Add references
references are for CNNs, approximation capabilities of NNs, and Experience Replay

@@
   Reinforcement learning is a subarea of machine learning inspired by neuroscience, where learners have to select actions with
-  regards to the environement state and receive a reward accordingly. Each agent's objective is to maximize its cumulative reward across its whole lifetime,
+  regards to the environment state and receive a reward accordingly. Each agent's objective is to maximize its cumulative reward across its whole lifetime,
   subdivided into episodes. In the traditional framework of reinforcement learning and more specifically in Q-learning,
@@
   These techniques have revealed themselves to be able to beat human experts at playing Atari games
-  by learning only from visual features. Despite the current progress, learning to take complex decisions on the basis of high-dimensional visual data remains on
+  by learning only from visual features. Despite the current progress, learning how to take complex decisions on the basis of high-dimensional visual data remains an
   ongoing challenge. In particular, because of the instability of neural networks, we also explored state-of-the-art dueling architectures to enhance performance
@@
 Instead, one can design a learnable function that maps pixel values to a single row of this table, thus a vector of Q-values whose length is the size
-of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article \citep{Mnih2015}.
-A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers. One characteristic of CNNs is the
+of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{Mnih2015}, since feedforward Neural Networks
+are known to be universal function approximators~\citep{hornik1991approximation}.
+A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers~\citep{lecun1998gradient}. One characteristic of CNNs is the
 presence of convolutional filters that maps pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
 allows the convolutional layer to capture the local information of the pixels, contrary to what dense (fully-connected) layers do. This procedure is greatly
-inspired by the notion of receptive field introduced by Hubel and Wiesel \citep{Hubel1962}.
+inspired by the notion of receptive field introduced by Hubel and Wiesel~\citep{Hubel1962}.
 Let $Q(s, a; \theta_i)$ be the Q-value approximation function represented by the neural network, parameterized by its weights $\theta$ at time step $i$.
@@
 To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse to the experience
-replay algorithm. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
-We compared this technique to a more efficient approach called prioritized experience replay \citep{DBLP:journals/corr/SchaulQAS15}.
+replay algorithm ~\citep{adam2012experience}. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
+We compared this technique to a more efficient approach called prioritized experience replay~\citep{DBLP:journals/corr/SchaulQAS15}.
 To emphasize the importance of the memory replay mechanism, we compared the results obtained using prioritized experience replay, regular
@@
 \[
-    P = 100 \times (score - RAS) / (HS - RAS)
+    P = 100 \times (\text{score} - RAS) / (HS - RAS)
 \]
@@
 
-We developed our models with the support of the Arcade Learning Environment \citep{bellemare13arcade}. This gives the opportunity to either
+We developed our models with the support of the Arcade Learning Environment~\citep{bellemare13arcade}. This gives the opportunity to either
 retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
@@
 \[
-    luminance = 0.299 \times R + 0.587 \times G + 0.114 \times B
+    \text{luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B
 \]
@@
 
-  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop} \citep{wattenberg2016how}
+  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop}~\citep{wattenberg2016how}
 
@@
 
-  \todo{Amélioration: Double dueling DQN} \citep{DBLP:journals/corr/WangFL15}
+  \todo{Amélioration: Double dueling DQN}~\citep{DBLP:journals/corr/WangFL15}
 

### 2017-12-27 agent.py
Fix verification of checkpoint existence, auto-save every 1000 steps

@@ -38,12 +38,14 @@ class Agent:
     
     def load_session(self):
         save_file = path.join(Parameters.SESSION_SAVE_DIRECTORY, Parameters.SESSION_SAVE_FILENAME)
-        if path.exists(save_file) and path.isfile(save_file):
+        if path.exists(Parameters.SESSION_SAVE_DIRECTORY):
                 # restore from a previously saved session
+                print("Loading session from", save_file)
                 tf_saver = tf.train.Saver()
                 tf_saver.restore(self.tf_session, save_file)
         else:
                 # initialize from scratch
+                print("Loading new session")
                 self.tf_session.run(tf.global_variables_initializer())
                 
     
@@ -53,6 +55,7 @@ class Agent:
                 makedirs(Parameters.SESSION_SAVE_DIRECTORY)
         tf_saver = tf.train.Saver()
         tf_saver.save(self.tf_session, save_file)
+        print("Saved session to", save_file)
 
 
     def train(self):
@@ -85,7 +88,7 @@ class Agent:
                 
                 self.step += 1
                 
-                if self.step % 10000 == 0:
+                if self.step % 1000 == 0:
                         self.save_session()
         
         self.save_session()

### 2017-12-27 parameters.py

@@ -41,7 +41,7 @@ class Parameters:
             data = json.load(f)
             for key in data.keys():
                 if type(data[key]) == dict:
-                    if key == "SESSION_SAVE_FILENAME":
-                        # make session filename specific to the game being played
-                        data[key]["value"] = data[key]["value"].replace(".", "_"+Parameters.GAME+".")
+                    if key in ("SESSION_SAVE_FILENAME", "SESSION_SAVE_DIRECTORY"):
+                        # make session path specific to the game being played
+                        data[key]["value"] = data[key]["value"] + Parameters.GAME
                     Parameters.add_attr(key, data[key]["value"])

### 2017-12-27 agent.py
Remove debug outputs, do not sleep between steps

@@ -156,7 +156,6 @@ class Agent:
         dt = self.step - Parameters.REPLAY_START_SIZE
         df = Parameters.FINAL_EXPLORATION_FRAME - Parameters.REPLAY_START_SIZE
         eps = Parameters.INITIAL_EXPLORATION - ((dt / df) * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
-        eps = 0
         if random.random() < eps:
             # take a random action
             action = np.random.randint(0, Parameters.ACTION_SPACE, size=1)[0]
@@ -165,7 +164,6 @@ class Agent:
             input_shape = (1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)
             dqn_input = self.environment.get_input().reshape(input_shape)
             q_print = self.tf_session.run(self.dqn.q_values, {self.dqn_input: dqn_input})
-            print("Q-values: ", q_print, "  at step : ", self.step)
             action = self.tf_session.run(self.dqn.smartest_action, {self.dqn_input: dqn_input})
         
         return(action)

### 2017-12-27 parameters.py

@@ -14,7 +14,7 @@ class Parameters:
     GAME = GAMES.SPACE_INVADERS
     ACTION_SPACE = GAMES.get_action_space(GAME)
     DISPLAY = True
-    SLEEP_BETWEEN_STEPS = True
+    SLEEP_BETWEEN_STEPS = False
 
     @staticmethod
     def add_attr(name, value):

### 2017-12-27 article/article.tex
[Article] Descripe the CNN structure

@@
 Universit\'{e} Libre de Bruxelles \\
-apassemi@ulb.ac.be}
+\{Cedric.Simar, apassemi, sgueniff, robpetit\}@ulb.ac.be}
 
@@
 \[
-  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k | s_t = s, a_t = a, \pi\right].
+  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k \; \big| \; s_t = s, a_t = a, \pi\right].
 \]
@@
 of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{Mnih2015}, since feedforward Neural Networks
-are known to be universal function approximators~\citep{hornik1991approximation}.
+are known to be universal function approximators (Theorem 2 in~\cite{hornik1991approximation}).
 A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers~\citep{lecun1998gradient}. One characteristic of CNNs is the
@@
 \[
-    \text{luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B
+    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B
 \]
@@
 
-\todo{Suite: CNN architecture}
+The Neural Network is composed of three CNNs followed by a fully connected layer, leading to the final fully connected layer which is the output.
+Between each of these units is a non-linear unit: a Rectifier Linear Unit (ReLU) \citep{krizhevsky2012imagenet}, but an Exponential Linear Unit (ELU) could
+also be used since it has been shown to be more performant in classification and decision tasks~\citep{DBLP:journals/corr/ClevertUH15}, or even a
+Flexible Rectified Linear Unit~\citep{qiu2017flexible}.
+
+The first CNN outputs 32 $20 \times 20$ images from a $84 \times 84$ image, using $8 \times 8$ filters and a stride of $4$. The second one outputs
+64 $9 \times 9$ images from the previous 32, using filters of size $4 \times 4$ and a stride of $2$; and the third and last CNN outputs performs
+64 $3 \times 3$ filters with stride $1$ on the 64 subimages of the previous layer to output $7 \times 7$ images.
+
+These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully connected layer with 512 neurons,
+which is then connected to the last layer (again fully connected) which corresponds to the output with the same size as the action space (6 in the
+case of Space Invaders).
 

### 2017-12-27 article/article.tex
[Article] Add details on 'why a NN' and on weights update

@@
 \usepackage{amssymb}
+\usepackage{commath}
 
 \usepackage{hyperref}
+\usepackage{flushend}
 
@@
 cumulative reward. More formally, this cumulative calculation starts from present time $t$ and is computed as follows:
-\[
-  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k \; \big| \; s_t = s, a_t = a, \pi\right].
-\]
+\begin{equation}
+  Q^{*}(s, a) = \max_{\pi} \E\left[ \sum_{k \geq 0}r_{t+k}\gamma^k \; \big| \; s_t = s, a_t = a; \pi\right].
+\end{equation}
 where $\pi = P(a | s)$ is a behavior policy, $t$ is the current step, $r_t$ is the reward for step $t$, $a_t$ is the action selected at time $t$,
@@
 Instead, one can design a learnable function that maps pixel values to a single row of this table, thus a vector of Q-values whose length is the size
-of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{Mnih2015}, since feedforward Neural Networks
+of the action space. For that purpose we used a Deep Q-Network (DQN) to approximate this function, as in the original DQN article~\citep{mnih2015human}, since feedforward Neural Networks
 are known to be universal function approximators (Theorem 2 in~\cite{hornik1991approximation}).
@@
 refined over time. The deep neural network incorporates minimal prior knowledge about the environment: the only assumption lies in the number of
-outputs in the last hidden layer, which has been set as the number of actions. In every Atari game, the action space is discrete and finite: this allows
-us to determine the output shape of the neural network depending on the game itself. The number of actions made available by the environment
+outputs in the last hidden layer, which has been set as the number of actions. In every Atari game, the action space is discrete, finite and constant over time:
+this allows us to determine the output shape of the neural network depending on the game itself. The number of actions made available by the environment
 cannot be inferred by the DQN during training phase.
@@
 random action selection) using a normalized performance metric. The latter is formalized as below:
-\[
-    P = 100 \times (\text{score} - RAS) / (HS - RAS)
-\]
+\begin{equation}
+    P = 100 \times (\text{score} - RAS) / (HS - RAS),
+\end{equation}
 where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
@@
 
-Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}
+Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}.
 
@@
 in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
-Then the sample size has been reduced by a factor of three by mapping RBG values to the pixel luminance, using the following formula:
-\[
-    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B
-\]
+Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance, using the following formula:
+\begin{equation}
+    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B.
+\end{equation}
+
 After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
 $84 \times 84$ pixel grid.
+
+As estimating the whole $Q^*$ function is practically not performable, $Q^*$ is estimated by a non-linear function depending on $\theta$ such that:
+\begin{equation}
+	\forall s \in S : \forall a \in A(s) : Q(s, a; \theta) \simeq Q^*(s, a).
+\end{equation}
+
+In this situation, the non-linear function is the neural network, and $\theta$ represents its weights.
+
+The interest of using a neural network is to be able to update the weights such that at each time step $t$, the weights are $\theta_t$, and:
+\begin{equation}
+	\forall s \in S : \forall a \in A(s) : Q(s, a; \theta_t) \xrightarrow[t \to +\infty]{} Q^*(s, a).
+\end{equation}
+
+Therefore, the NN is trained by minimizing the Huber loss~\citep{huber1964robust} with $\delta=1$ of the Bellman equation (by a -- stochastic -- gradient descent).
+The Huber loss $H_\delta$, parameterized by $\delta \in {\mathbb R^*}^+$, is a continuous function (also $C^1$ but not $C^2$) defined by the MSE
+on a small neighbourhood of $0$, and by the absolute value outside this neighbourhood:
+\begin{equation}
+	H_\delta : \mathbb R \to \mathbb R : x \mapsto \begin{cases}\frac 12x^2                                &\text{ if } \abs x \leq \delta,\\
+	                                                            \delta\left(\abs x - \frac 12\delta\right) &\text{ otherwise.}\end{cases}
+\end{equation}
+
+The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and has (\todo {probably}) been shown to increase
+stability.~\todo{check if this has a proper reference or drop it}
+
+\subsection{Neural Network architecture}
 
@@
 case of Space Invaders).
+
+\subsection{Update of the weights}
+
+In order to reduce oscillation in $Q$ estimations, two separate neural networks have been used: the first one, denoted $Q$ and called the \textit{action-value}
+network is the one used in action prediction, and the second one denotes $\hat Q$ and called the \textit{target action-value} network is used when
+updating the weights. The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of $\hat Q$ are denoted by $\theta^-$.
+
+Every 4 steps, at time $t$, a minibatch of 32 past experiences $\left\{(s_{t_j}, a_{t_j}, r_{t_j}, s_{t_j+1})\right\}_{j=1}^{32}$ is selected uniformly from the
+memory in order to perform a Stochastic Gradient Descent (SGD), meaning that the average of the gradients of the loss function on each experience is used to update
+the weights. As mentioned above, the loss function is a Huber loss function and is defined as:
+\begin{equation}
+	H_1\left(y_j - Q(s_{t_j}, a_{t_j}; \theta)\right),
+\end{equation}
+with $y_j$ set either to $r_j$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
+$r_j + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.
+
+The $\hat Q$ target action-value network is then copied back from the action value network $Q$ every 10,000 steps.
 
@@
 
+%\newpage
 \footnotesize

### 2017-12-28 src/dddqn.py
DDDDDDDDDDDQN

@@ -1,0 +1,110 @@
+# -*- coding: utf-8 -*-
+# dddqn.py : Double dueling deep Q-network
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
+
+from dqn import DQN
+from tf_decorator import *
+from parameters import Parameters
+
+import tensorflow as tf
+from tensorflow.python.ops import array_ops as tf_array_ops
+
+
+class DDDQN(DQN):
+
+    def __init__(self, state):
+        DQN.__init__(self, state)
+
+    @define_scope
+    def q_values(self):
+        """
+        References
+        ----------
+        1)  https://arxiv.org/pdf/1511.06581.pdf
+            Dueling Network Architectures for Deep Reinforcement Learning
+            Wang et al.
+        2)  https://github.com/awjuliani/DeepRL-Agents/blob/master/Double-Dueling-DQN.ipynb
+            Simple Reinforcement Learning with Tensorflow Part 4: Deep Q-Networks and Beyond
+        """
+        
+        """
+        [Article] The input to the neural network (the state) consists of an 84 x 84 x 4 image 
+        produced by the preprocessing map "phi"
+        """
+        # reshape input to 4d tensor [batch, height, width, channels]
+        input = tf.reshape(self.state, [-1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH])
+
+        # convolutional layer 1
+        """
+        [Article] The first hidden layer convolves 32 filters of 8 x 8 with stride 4 with the
+        input image and applies a rectifier nonlinearity
+        """
+        W_conv1 = self.weight_variable([8, 8, Parameters.AGENT_HISTORY_LENGTH, 32])
+        b_conv1 = self.bias_variable([32])
+        conv1 = tf.nn.conv2d(input, W_conv1, strides=[1, 4, 4, 1], padding='VALID') # would 'SAME' also work ?
+        h_conv1 = tf.nn.relu(conv1 + b_conv1)
+        
+        # output of conv 1 is of shape [-1 x 20 x 20 x 32]
+        
+        # convolutional layer 2
+        """
+        [Article] The second hidden layer convolves 64 filters of 4 x 4 with stride 2, 
+        again followed by a rectifier nonlinearity
+        """
[…]

### 2017-12-28 src/agent.py
DDDDDDDDDDDQN

@@ -7,6 +7,7 @@ from environment import Environment
 from memory import Memory
 from parameters import Parameters
 from dqn import DQN
+from dddqn import DDDQN
 from os import path, makedirs
 
 import random
@@ -26,7 +27,7 @@ class Agent:
         
         # initialize the DQN and target DQN (with respective placeholders)
         self.dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "DQN_input")
-        self.dqn = DQN(self.dqn_input)
+        self.dqn = DDDQN(self.dqn_input)
 
         self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "target_DQN_input")
         self.target_dqn = DQN(self.target_dqn_input)

### 2017-12-28 src/memory.py
Storing long-term experience on HDD

@@ -1,18 +1,30 @@
+# -*- coding: utf-8 -*-
+# memory.py : Agent's short term and long term memory
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
 from parameters import Parameters
 import numpy as np
 
+
 class Memory:
 
-    def __init__(self):
+    def __init__(self, destination="mem.dat"):
+        """
+        :param destination: str
+            Path to the file where the long-term experience must be stored
+            Note: Files cannot be larger than 2 Gb in 32-bit architectures
+        """
 
-        self.memory_size = Parameters.REPLAY_MEMORY_SIZE
+        self.memory_filepath = destination
+
+        self.memory_size = Parameters.LONG_TERM_MEMORY_SIZE
         self.current_memory_index = 0
         self.memory_usage = 0
 
         self.actions = np.empty(self.memory_size, dtype=np.uint8)
         self.rewards = np.empty(self.memory_size, dtype=np.integer)
-        self.screens = np.empty((self.memory_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH), dtype=np.float16)
+        screens_shape = (self.memory_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH)
+        self.screens = np.memmap(self.memory_filepath, mode="w+", shape=screens_shape, dtype=np.float16)
         self.terminals = np.empty(self.memory_size, dtype=np.bool)
 
         self.minibatch_size = Parameters.MINIBATCH_SIZE

### 2017-12-28 src/main.py
Create a new dev set of hyper-parameters for convenience

@@ -9,7 +9,7 @@ from parameters import Parameters
 
 if __name__ == "__main__":
     
-    Parameters.load("parameters.json")
+    Parameters.load("parameters/dev.json")
     environment = Environment()
     agent = Agent(environment)
     

### 2017-12-28 article/article.tex
[Article] add explanation on reward clipping

@@
 \documentclass[letterpaper]{article}
-\usepackage{natbib,alifexi}
+
+\usepackage[utf8]{inputenc}
+\usepackage[sort, colon]{natbib}
+\usepackage{alifexi}
 \usepackage{color}
@@
 
-\usepackage{hyperref}
+\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
 \usepackage{flushend}
@@
 
+The huge breakout effect of the DQN was that it has shown to be a first step towards general artificial intelligence~\citep{togelius2015ai}, meaning an
+algorithm that is not only good at learning one particular task but several of them. Even though the DQN needs to be trained for about a million steps
+on each Atari game, it gets good at most of them and better than human at 22 of the 49 experimented~\citep{mnih2015human}.
+
 Since we are handling visual features, we used screen frames as environment state representations. Each frame is a (210 x 160) color image sampled at
@@
 $84 \times 84$ pixel grid.
+
+The perceived rewards of the agent have all been reduced to be either $-1$ for a negative reward, $0$ for a session with no reward, and $+1$
+for a positive reward, i.e. at any time $t$: $r_t \in \{-1, 0, +1\}$. No intermediate or larger reward is available to the learning algorithm.
+This allows to learn without any previous knowledge on the particular game the DQN is trained on. Yet some rewards can be highly different (in Pac-Man likes,
+eating a cherry gives 100 points, a banana gives 5000, and eating a ghost gives $2^n \times 100$ with $n$ the number of eaten ghosts in a single run).
+This could be improved by adaptive normalization~\citep{van2016learning}.
 

### 2017-12-28 src/agent.py
Xavier initialization in DDDQN

@@ -164,7 +164,10 @@ class Agent:
             # take a smart action
             input_shape = (1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)
             dqn_input = self.environment.get_input().reshape(input_shape)
-            q_print = self.tf_session.run(self.dqn.q_values, {self.dqn_input: dqn_input})
+
+            q_values = self.tf_session.run(self.dqn.q_values, {self.dqn_input: dqn_input})
+            # Plotter.add_q_values_at_t(q_values)
+
             action = self.tf_session.run(self.dqn.smartest_action, {self.dqn_input: dqn_input})
         
         return(action)

### 2017-12-28 src/dddqn.py

@@ -78,8 +78,8 @@ class DDDQN(DQN):
         sv_stream = tf.contrib.layers.flatten(sv_stream)
 
         # Add one fully connected linear layer per stream
-        W_apa = self.weight_variable([1568, Parameters.ACTION_SPACE])
-        W_sv = self.weight_variable([1568, 1])
+        W_apa = self.weight_variable([1568, Parameters.ACTION_SPACE], method="xavier")
+        W_sv = self.weight_variable([1568, 1], method="xavier")
         fc_apa = tf.matmul(apa_stream, W_apa)
         fc_sv = tf.matmul(sv_stream, W_sv)
         

### 2017-12-28 src/dqn.py

@@ -186,12 +186,18 @@ class DQN:
         return(self.mean_error)
 
 
-    def weight_variable(self, shape):
+    def weight_variable(self, shape, method="normal"):
         """
-        Initialize weight variable randomly using a truncated normal distribution
-        of mean = 0 and standard deviation of 0.02
+        Initialize weight variable randomly using one of the two following methods:
+            - A truncated normal distribution of mean = 0 and standard deviation of 0.02
+            - A xavier initialization, where var(W) = 1 / n_in
         """
-        weight_var = tf.truncated_normal(shape, mean = 0, stddev=0.02)
+        assert(method in ["normal", "xavier"])
+        if method == "normal":
+            weight_var = tf.truncated_normal(shape, mean = 0, stddev=0.02)
+        else:
+            xavier_initializer = tf.contrib.layers.xavier_initializer()
+            weight_var = xavier_initializer(list(shape))
         return(tf.Variable(weight_var))
 
 
@@ -235,10 +241,4 @@ class DQN:
         else:
             print("Thou shall only assign learning parameters!")
 
-    
-    def save_learning_parameters(self):
-        print("TODO")
-
-    def load_learning_parameters(self):
-        print("TODO")
 

### 2017-12-28 src/agent.py
Plot avg q-value per batch and Total score per episode

@@ -6,10 +6,11 @@ from utils import reward_clipper
 from environment import Environment
 from memory import Memory
 from parameters import Parameters
+from plot import Plotter
 from dqn import DQN
 from dddqn import DDDQN
-from os import path, makedirs
 
+from os import path, makedirs
 import random
 import time
 
@@ -21,16 +22,17 @@ class Agent:
 
     def __init__(self, environment):
         
+        self.action_space = Parameters.GAMES.get_action_space(Parameters.GAME)
         self.environment = environment
         self.memory = Memory()
         self.step = 0
         
         # initialize the DQN and target DQN (with respective placeholders)
         self.dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "DQN_input")
-        self.dqn = DDDQN(self.dqn_input)
+        self.dqn = DDDQN(self.dqn_input, self.action_space)
 
         self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "target_DQN_input")
-        self.target_dqn = DQN(self.target_dqn_input)
+        self.target_dqn = DQN(self.target_dqn_input, self.action_space)
 
         # initialize the tensorflow session and variables
         self.tf_session = tf.Session()
@@ -56,6 +58,7 @@ class Agent:
                 makedirs(Parameters.SESSION_SAVE_DIRECTORY)
         tf_saver = tf.train.Saver()
         tf_saver.save(self.tf_session, save_file)
+        Plotter.save("out")
         print("Saved session to", save_file)
 
 
@@ -101,6 +104,8 @@ class Agent:
         Apply Q-learning updates, or minibatch updates, to samples of experience,
         (s, a, r, s') ~ U(D), drawn at random from the pool of stored samples.
         """
+        
+        Plotter.notify_batch() # Notify the plotter that the next data points will correspond to a new batch
 
         if(self.memory.get_usage() > Parameters.AGENT_HISTORY_LENGTH):
             
@@ -159,14 +164,14 @@ class Agent:
         eps = Parameters.INITIAL_EXPLORATION - ((dt / df) * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
         if random.random() < eps:
[…]

### 2017-12-28 src/dddqn.py

@@ -12,8 +12,8 @@ from tensorflow.python.ops import array_ops as tf_array_ops
 
 class DDDQN(DQN):
 
-    def __init__(self, state):
-        DQN.__init__(self, state)
+    def __init__(self, state, action_space):
+        DQN.__init__(self, state, action_space)
 
     @define_scope
     def q_values(self):
@@ -78,7 +78,7 @@ class DDDQN(DQN):
         sv_stream = tf.contrib.layers.flatten(sv_stream)
 
         # Add one fully connected linear layer per stream
-        W_apa = self.weight_variable([1568, Parameters.ACTION_SPACE], method="xavier")
+        W_apa = self.weight_variable([1568, self.action_space], method="xavier")
         W_sv = self.weight_variable([1568, 1], method="xavier")
         fc_apa = tf.matmul(apa_stream, W_apa)
         fc_sv = tf.matmul(sv_stream, W_sv)

### 2017-12-28 src/dqn.py

@@ -11,10 +11,13 @@ from tensorflow.python.ops import array_ops as tf_array_ops
 
 class DQN:
 
-    def __init__(self, state):
+    def __init__(self, state, action_space):
 
         # receiving state placeholder
         self.state = state
+
+        # set action space size
+        self.action_space = action_space
 
         # weights and biases dictionary
         self.learning_parameters = {}
@@ -96,11 +99,11 @@ class DQN:
         h_fc1 = tf.nn.relu(fc1 + b_fc1)
 
         # fully connected layer 2 (output layer)
-        W_fc2 = self.weight_variable([512, Parameters.ACTION_SPACE])
-        b_fc2 = self.bias_variable([Parameters.ACTION_SPACE])
+        W_fc2 = self.weight_variable([512, self.action_space])
+        b_fc2 = self.bias_variable([self.action_space])
         fc2 = tf.matmul(h_fc1, W_fc2)
 
-        # network output is of shape (1, Parameters.ACTION_SPACE)
+        # network output is of shape (1, self.action_space)
         """
         [Article] We use an architecture in which there is a separate output unit 
         for each possible action [ = one-hot encoding ], and only the state representation 
@@ -158,7 +161,7 @@ class DQN:
         self.action = tf.placeholder(tf.int64, [None], name="action")
 
         # convert the action to one-hot representation in order to compute the error
-        action_one_hot = tf.one_hot(self.action, Parameters.ACTION_SPACE, on_value=1, off_value=0, name="action_one_hot")
+        action_one_hot = tf.one_hot(self.action, self.action_space, on_value=1, off_value=0, name="action_one_hot")
         
         self.q_acted = tf.reduce_sum(self.q_values * tf.cast(action_one_hot, tf.float32), axis=1, name="q_acted")
         

### 2017-12-28 src/environment.py

@@ -4,6 +4,7 @@
 
 from history import FramesHistory
 from parameters import Parameters
+from plot import Plotter
 import utils
 
 import gym
@@ -40,7 +41,8 @@ class Environment:
         self.take_action(first_action)
         self.render()
         self.reward = 0
-        self.terminal = False # True? -> Parameter ?
+        self.episode_score = 0
+        self.terminal = False
 
         self.initialize_screens_history()
 
@@ -99,6 +101,7 @@ class Environment:
             
             self.take_action(action)
             cumulated_reward += self.reward
+            self.episode_score += self.reward
 
             """
             [Article] For games where there is a life counter, the Atari
@@ -108,6 +111,9 @@ class Environment:
             if self.lives < lives_before_action:
                 cumulated_reward += Parameters.NEGATIVE_REWARD
                 self.terminal = True
+
+                Plotter.add_episode_score(self.episode_score)
+                self.episode_score = 0
             
             skipped += 1
 

### 2017-12-28 src/games.py

@@ -9,7 +9,6 @@ class Games:
 
     def __init__(self):   
 
-        # Cool (easy) Atari environments available
         self.ASTEROIDS = "Asteroids-v0"
         self.PONG = "Pong-v0"
         self.SPACE_INVADERS = "SpaceInvaders-v0"
@@ -17,8 +16,6 @@ class Games:
 
         self.available_games_list = [self.ASTEROIDS, self.PONG, self.SPACE_INVADERS, self.TENNIS]
         self.action_space = {}
-        self.define_action_spaces()
-
 
     def define_action_spaces(self):
         
@@ -32,6 +29,8 @@ class Games:
             assert(type(actions) == gym.spaces.discrete.Discrete)
 
     def get_action_space(self, game):
+        if len(self.action_space.values()) < len(self.available_games_list):
+            self.define_action_spaces()
         return(self.action_space[game])
 
         

### 2017-12-28 src/main.py

@@ -2,17 +2,34 @@
 # main.py 
 # author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
 
+from plot import Plotter
 from agent import Agent
 from environment import Environment
 from parameters import Parameters
 
+import argparse
 
-if __name__ == "__main__":
-    
+
+def train():
+    Plotter.load("out")
     Parameters.load("parameters/dev.json")
     environment = Environment()
     agent = Agent(environment)
-    
     agent.train()
 
-    
+def plot_figures():
+    folder = "out/"
+    Plotter.load(folder)
+    Plotter.save_plots(folder)
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser()
+    group = parser.add_mutually_exclusive_group(required=True)
+    group.add_argument('--train', action='store_true', help='Make the agent learn')
+    group.add_argument('--plot', action='store_true', help='Plot the results obtained during training')
+
+    args = parser.parse_args()
+    if args.train:
+        train()
+    else:
+        plot_figures()

### 2017-12-28 src/memory.py

@@ -110,3 +110,9 @@ class Memory:
     
     def get_usage(self):
         return(self.memory_usage)
+
+
+class PrioritizedMemory(Memory):
+    def __init__(self, *args, **kwargs):
+        Memory.__init__(self, *args, **kwargs)
+        #  TODO : after dodo :'( 

### 2017-12-28 src/parameters.py

@@ -12,7 +12,6 @@ class Parameters:
     # environment settings
     GAMES = Games()
     GAME = GAMES.SPACE_INVADERS
-    ACTION_SPACE = GAMES.get_action_space(GAME)
     DISPLAY = True
     SLEEP_BETWEEN_STEPS = False
 

### 2017-12-28 src/plot.py

@@ -1,0 +1,91 @@
+# -*- coding: utf-8 -*-
+# plot.py : Graphs, figures and manifold learning
+# author : Robin Petit, Stanislas Gueniffey, Cedric Simar, Antoine Passemiers
+
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def warn_if_empty(func):
+    def new_func(*args, **kwargs):
+        try:
+            func(*args, **kwargs)
+        except IndexError:
+            print("[Warning] Could not save figure in %s: empty sequence" % func.__name__)
+    return new_func
+
+
+class Plotter:
+
+    Q_VALUES = list()
+    Q_VALUES_CURRENT_BATCH = list()
+    SCORES_PER_EPISODE = list()
+
+    @staticmethod
+    def add_episode_score(score):
+        Plotter.SCORES_PER_EPISODE.append(score)
+
+    @staticmethod
+    def add_q_values_at_t(q_values):
+        Plotter.Q_VALUES_CURRENT_BATCH.append(np.squeeze(q_values))
+
+    @staticmethod
+    def notify_batch():
+        if len(Plotter.Q_VALUES_CURRENT_BATCH) > 0:
+            avg_q_values = np.nan_to_num(np.asarray(Plotter.Q_VALUES_CURRENT_BATCH)).mean(axis = 0)
+            Plotter.Q_VALUES.append(avg_q_values)
+            Plotter.Q_VALUES_CURRENT_BATCH = list()
+
+    @staticmethod
+    @warn_if_empty
+    def plot_e_scores():
+        plt.plot(Plotter.SCORES_PER_EPISODE)
+        plt.xlabel("Episode")
+        plt.ylabel("Average score per episode")
+
+    @staticmethod
+    @warn_if_empty
+    def plot_q_values():
+        avg_q_values = np.asarray(Plotter.Q_VALUES).mean(axis=1)
+        plt.plot(avg_q_values)
+        plt.xlabel("Number of batches")
+        plt.ylabel("Average Q-values")
+
+    @staticmethod
+    def save_plots(folder):
+        if not os.path.exists(folder):
+            os.makedirs(folder)
+        
+        Plotter.plot_e_scores()
+        plt.savefig(os.path.join(folder, "e_scores.png"))
+        plt.clf(); plt.close()
+
+        Plotter.plot_q_values()
+        plt.savefig(os.path.join(folder, "q_values.png"))
+        plt.clf(); plt.close()
+
+    @staticmethod
+    def load(folder):
[…]

### 2017-12-28 src/agent.py
Smoothed plots, and option for resetting plots

@@ -4,7 +4,7 @@
 
 from utils import reward_clipper
 from environment import Environment
-from memory import Memory
+from memory import Memory, PrioritizedMemory
 from parameters import Parameters
 from plot import Plotter
 from dqn import DQN
@@ -24,7 +24,7 @@ class Agent:
         
         self.action_space = Parameters.GAMES.get_action_space(Parameters.GAME)
         self.environment = environment
-        self.memory = Memory()
+        self.memory = PrioritizedMemory()
         self.step = 0
         
         # initialize the DQN and target DQN (with respective placeholders)
@@ -32,7 +32,7 @@ class Agent:
         self.dqn = DDDQN(self.dqn_input, self.action_space)
 
         self.target_dqn_input = tf.placeholder(tf.float32, [None, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH], name = "target_DQN_input")
-        self.target_dqn = DQN(self.target_dqn_input, self.action_space)
+        self.target_dqn = DDDQN(self.target_dqn_input, self.action_space)
 
         # initialize the tensorflow session and variables
         self.tf_session = tf.Session()
@@ -93,7 +93,7 @@ class Agent:
                 self.step += 1
                 
                 if self.step % 1000 == 0:
-                        self.save_session()
+                    self.save_session()
         
         self.save_session()
             
@@ -136,7 +136,7 @@ class Agent:
         
         # update agent's memory and environment's history
         self.environment.add_current_screen_to_history()
-        self.memory.add(screen, action, reward_clipper(reward), terminal)
+        self.memory.add(screen, action, reward_clipper(reward), terminal,)
 
         # if we started learning
         if(self.step > Parameters.REPLAY_START_SIZE):
@@ -169,11 +169,10 @@ class Agent:
             # take a smart action
             input_shape = (1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)
             dqn_input = self.environment.get_input().reshape(input_shape)
+            action = self.tf_session.run(self.dqn.smartest_action, {self.dqn_input: dqn_input})
 
             q_values = self.tf_session.run(self.dqn.q_values, {self.dqn_input: dqn_input})
             Plotter.add_q_values_at_t(q_values)
-
-            action = self.tf_session.run(self.dqn.smartest_action, {self.dqn_input: dqn_input})
         
         return(action)
 

### 2017-12-28 src/main.py

@@ -9,27 +9,31 @@ from parameters import Parameters
 
 import argparse
 
+OUT_FOLDER = "out"
 
 def train():
-    Plotter.load("out")
+    Plotter.load(OUT_FOLDER)
     Parameters.load("parameters/dev.json")
     environment = Environment()
     agent = Agent(environment)
     agent.train()
 
 def plot_figures():
-    folder = "out/"
-    Plotter.load(folder)
-    Plotter.save_plots(folder)
+    Plotter.load(OUT_FOLDER)
+    Plotter.save_plots(OUT_FOLDER)
 
 if __name__ == "__main__":
     parser = argparse.ArgumentParser()
     group = parser.add_mutually_exclusive_group(required=True)
     group.add_argument('--train', action='store_true', help='Make the agent learn')
     group.add_argument('--plot', action='store_true', help='Plot the results obtained during training')
+    group.add_argument('--reset-plot', action='store_true', help='Delete results obtained during training')
 
     args = parser.parse_args()
     if args.train:
         train()
-    else:
+    elif args.plot:
         plot_figures()
+    elif args.reset_plot:
+        Plotter.reset(OUT_FOLDER)
+        print("Training results deleted from folder %s" % OUT_FOLDER)

### 2017-12-28 src/memory.py

@@ -35,6 +35,7 @@ class Memory:
 
 
     def add(self, screen, action, reward, terminal):
+        # q_estimates are not used by the Memory, but by the PrioritizedMemory
 
         self.actions[self.current_memory_index] = action
         self.rewards[self.current_memory_index] = reward
@@ -50,7 +51,7 @@ class Memory:
         state = None
 
         if(not self.memory_usage):
-            print("Memory is empty")
+            print("[Warning] Memory is empty")
         else:
 
             state_index = state_index % self.memory_usage
@@ -113,6 +114,27 @@ class Memory:
 
 
 class PrioritizedMemory(Memory):
-    def __init__(self, *args, **kwargs):
-        Memory.__init__(self, *args, **kwargs)
-        #  TODO : after dodo :'( 
+    """
+    References
+    ----------
+    https://arxiv.org/pdf/1511.05952.pdf
+    Prioritized experience replay
+    Schaul et al.
+    """
+    def __init__(self, alpha=1.0, destination="mem.dat"):
+        """
+        :param alpha: float
+            Degree of prioritization
+            In the case of uniform experience replay, alpha = 0
+        :param destination: str
+            Path to the file where the long-term experience must be stored
+            Note: Files cannot be larger than 2 Gb in 32-bit architectures
+        """
+        Memory.__init__(self, destination=destination)
+        self.alpha = alpha
+        self.priorities = np.full(self.memory_size, fill_value=np.inf, dtype=np.float32)
+
+        # TODO: soon
+
+
+

### 2017-12-28 src/plot.py

@@ -15,6 +15,11 @@ def warn_if_empty(func):
             print("[Warning] Could not save figure in %s: empty sequence" % func.__name__)
     return new_func
 
+def moving_average(sequence, n=100):
+    ma = np.cumsum(sequence)
+    ma[n:] = ma[n:] - ma[:-n]
+    return ma[n-1:] / n
+
 
 class Plotter:
 
@@ -40,7 +45,8 @@ class Plotter:
     @staticmethod
     @warn_if_empty
     def plot_e_scores():
-        plt.plot(Plotter.SCORES_PER_EPISODE)
+        smoothed = moving_average(Plotter.SCORES_PER_EPISODE, n=50)
+        plt.plot(smoothed)
         plt.xlabel("Episode")
         plt.ylabel("Average score per episode")
 
@@ -48,7 +54,8 @@ class Plotter:
     @warn_if_empty
     def plot_q_values():
         avg_q_values = np.asarray(Plotter.Q_VALUES).mean(axis=1)
-        plt.plot(avg_q_values)
+        smoothed = moving_average(avg_q_values)
+        plt.plot(smoothed)
         plt.xlabel("Number of batches")
         plt.ylabel("Average Q-values")
 
@@ -85,6 +92,13 @@ class Plotter:
         q_values_path = os.path.join(folder, "q_values.npy")
         np.save(e_scores_path, np.asarray(Plotter.SCORES_PER_EPISODE))
         np.save(q_values_path, np.asarray(Plotter.Q_VALUES))
+    
+    @staticmethod
+    def reset(folder):
+        for name in os.listdir(folder):
+            filepath = os.path.join(folder, name)
+            if filepath.endswith(".png") or filepath.endswith(".npy"):
+                os.remove(filepath)
         
 
 class EmbeddingProjector:

### 2018-01-01 src/memory.py
get the whole batch when calling `Memory.sample_memory`

@@ -14,13 +14,22 @@ DEFAULT_SAVE_PATH = "memory-" + Parameters.GAME + '.shelf'
 
 class ShortTermMemory:
 
-    def __init__(self, mmap):
+    def __init__(self, mmap, memory):
         """
         :param mmap: np.memmap
             Memory map that holds all the experience samples.
             This map represents the long-term memory.
         """
+        self.parent_memory = memory
         self.long_term_mem = mmap
+        buffer_shape = list(mmap.shape)
+        buffer_shape[0] = Parameters.SHORT_TERM_MEMORY_SIZE
+        self.buffer = np.zeros(shape=buffer_shape)
+
+
+    def sample_random(self):
+        random_indices = np.random.choice(self.parent_memory.memory_usage, self.buffer.shape[0])
+        self.buffer = self.long_term_mem[random_indices, :]
 
 
 class Memory:
@@ -60,8 +69,17 @@ class Memory:
 
 
     def load_memory(self, path=DEFAULT_SAVE_PATH):
-        ret = os.path.exists(path)
-        if not ret:
+        try:
+            with shelve.open(path) as shelf:
+                self.current_memory_index = shelf["idx"]
+                self.memory_usage = shelf["mem usage"]
+                self.actions = shelf["actions"]
+                self.rewards = shelf["rewards"]
+                self.terminals = shelf["terminals"]
+                self.state_t = shelf["state t"]
+                self.state_t_plus_1 = shelf["state t+1"]
+            return True
+        except KeyError:
             self.current_memory_index = 0
             self.memory_usage = 0
             self.actions = np.empty(self.memory_size, dtype=np.uint8)
@@ -75,17 +93,7 @@ class Memory:
                 os.remove(self.memory_filepath+'.bak')
             except FileNotFoundError:
                 pass  # remove files if they exist so if they don't exist, just ignore
-        else:
-            with shelve.open(path) as shelf:
-                self.current_memory_index = shelf["idx"]
-                self.memory_usage = shelf["mem usage"]
-                self.actions = shelf["actions"]
-                self.rewards = shelf["rewards"]
-                self.terminals = shelf["terminals"]
-                self.state_t = shelf["state t"]
-                self.state_t_plus_1 = shelf["state t+1"]
-            print('Loaded memory from', path)
-        return ret
+            return False
 
 
     def add(self, screen, action, reward, terminal):
@@ -121,8 +129,12 @@ class Memory:
 
         return(state)
 
[…]

### 2018-01-01 src/agent.py
Enhance save timing display + add estimation of remaining time

@@ -11,6 +11,7 @@ from dqn import DQN
 from dddqn import DDDQN
 
 from os import path, makedirs
+from datetime import datetime, timedelta
 import random
 import time
 import sys
@@ -64,7 +65,8 @@ class Agent:
         self.tf_session = tf.Session()
         self.tf_saver = tf.train.Saver()
         self.load_session()
-        self.last_time = time.time()
+        self.initial_time = self.last_time = time.time()
+        self.initial_step = self.step
 
 
     def load_session(self):
@@ -82,14 +84,14 @@ class Agent:
 
     def save_session(self):
         time_at_start_save = time.time()
-        print('[Step {}  --  Took {:3.2f} s]'.format(self.step, time_at_start_save - self.last_time), end='\t')
+        print('{}: [Step {}k  --  Took {:3.2f} s]'.format(datetime.now(), self.step/1000, time_at_start_save - self.last_time), end=' ')
         self.last_time = time_at_start_save
 
         Parameters.CURRENT_STEP = self.step
         a = time.time()
         Parameters.update()
         b = time.time()
-        print('[{:3.2f}s for json]'.format(b-a), end='\t')
+        print('[{:3.2f}s for json]'.format(b-a), end=' ')
 
         save_file = path.join(Parameters.SESSION_SAVE_DIRECTORY, Parameters.SESSION_SAVE_FILENAME)
         if not path.exists(Parameters.SESSION_SAVE_DIRECTORY):
@@ -97,20 +99,22 @@ class Agent:
         a = time.time()
         self.tf_saver.save(self.tf_session, save_file)
         b = time.time()
-        print('[{:3.2f}s for tf]'.format(b-a), end='\t')
+        print('[{:3.2f}s for tf]'.format(b-a), end=' ')
 
         a = time.time()
         Plotter.save("out")
         b = time.time()
-        print('[{:3.2f}s for Plotter]'.format(b-a), end='\t')
+        print('[{:3.2f}s for Plotter]'.format(b-a), end=' ')
         a = time.time()
         self.memory.save_memory()
         b = time.time()
-        print('[{:3.2f}s for memory]'.format(b-a), end='\t')
+        print('[{:3.2f}s for memory]'.format(b-a), end=' ')
         post_save_time = time.time()
-        print('[Required {:3.2f}s to save all]'.format(post_save_time-time_at_start_save), end='\t')
+        print('[Required {:3.2f}s to save all]'.format(post_save_time-time_at_start_save), end=' ')
[…]

### 2018-01-01 article/article.tex

@@
 
+Figure~\ref{fig:t-SNE}\todo{Link the right figure} shows a 2-dimensional representation of the last fully connected hidden layer, i.e. a 512-dimensional
+vector. The method of \textit{projection} that was used is t-SNE~\citep{maaten2008visualizing}, which is part of the manifold learning algorithms, also
+called \textit{Nonlinear Dimensionality Reduction}~\citep{lee2007nonlinear}.
+
+The implementation of t-SNE that was used is the one from Scikit-Learn~\citep{scikit-learn}, the Python library.
+
   \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop}~\citep{wattenberg2016how}
@@
 
-%\newpage
+\newpage
 \footnotesize

### 2018-01-01 src/memory.py
Bug fix: retrieve all memory samples instead of just the 1st one

@@ -260,7 +260,7 @@ class PrioritizedMemory(Memory):
     def sample_memory(self, nb_samples=1):
         self.update_probs_and_weights()
         probs = self.sampling_probs[:self.memory_usage]
-        return np.random.choice(self.memory_usage, size=nb_samples, p=probs)[0]
+        return np.random.choice(self.memory_usage, size=nb_samples, p=probs)
 
     def update_probs_and_weights(self):
         probs = self.sampling_probs[:self.memory_usage]

### 2018-01-01 src/agent.py
Fix remaining time estimation

@@ -113,7 +113,7 @@ class Agent:
         print('[Required {:3.2f}s to save all]'.format(post_save_time-time_at_start_save), end=' ')
         self.last_time = post_save_time
         elapsed_time = time.time() - self.initial_time
-        remaining_seconds = elapsed_time*(Parameters.MAX_STEPS - self.step)/self.step
+        remaining_seconds = elapsed_time*(Parameters.MAX_STEPS - self.step)/(self.step - self.initial_step)
         print("eta: {}s".format((timedelta(seconds=remaining_seconds))))
 
 

### 2018-01-01 src/agent.py
Allow to play with pre-trained NN
use the parameter `--play` provided to `main.py`

@@ -214,7 +214,7 @@ class Agent:
         df = float(Parameters.FINAL_EXPLORATION_FRAME - Parameters.REPLAY_START_SIZE)
         return (dt / df)
 
-    def select_action(self):
+    def select_action(self, eps=None):
         """
         The agent uses its experience and expertise acquired
         through deep learning to make intelligent actions (sometimes)
@@ -222,7 +222,8 @@ class Agent:
 
         # compute epsilon at step t
         completion = self.get_learning_completion()
-        eps = Parameters.INITIAL_EXPLORATION - (completion * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
+		if eps is None:
+        	eps = Parameters.INITIAL_EXPLORATION - (completion * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
         if random.random() < eps:
             # take a random action
             action = np.random.randint(0, self.action_space, size=1)[0]
@@ -253,8 +254,15 @@ class Agent:
 
 
     def play(self):
-        print("To do")
-
-
-
+		self.environment.render(mode='human')
+        while True:
+            self.environment.terminal = False
+            while self.environment.get_lives():
+                action = self.select_action(eps=.1)
+                _, reward, done = self.environment.process_step(action)
+                if done:
+                    self.environment.reset()
+                    self.environment.terminal = False
+                time.sleep(1. / Parameters.FPS)
+                self.environment.render()
 

### 2018-01-01 src/main.py

@@ -34,6 +34,12 @@ def play_random(params_path):
     all_scores = agent.play()
     print('mean: ', np.mean(all_scores), '\tstd: ', np.std(all_scores))
 
+def play_pre_trained(params_path):
+    Parameters.load(params_path)
+    environment = Environment()
+    agent = Agent(environment)
+    agent.play()
+
 if __name__ == "__main__":
     parser = argparse.ArgumentParser()
     parser.add_argument('parameters_json', type=str, help='The path of the parameters to load (! Will be edited when saving!)')
@@ -42,6 +48,7 @@ if __name__ == "__main__":
     group.add_argument('--plot', action='store_true', help='Plot the results obtained during training')
     group.add_argument('--reset-plot', action='store_true', help='Delete results obtained during training')
     group.add_argument('--random', action='store_true', help='Play 500 games with random action selection and print the mean/std')
+	group.add_argument('--play', action='store_true', help='Play the game with pre-trained model')
 
     args = parser.parse_args()
     assert exists(args.parameters_json)
@@ -56,3 +63,5 @@ if __name__ == "__main__":
         print("Training results deleted from folder %s and Memory was removed from root." % OUT_FOLDER)
     elif args.random:
         play_random()
+    elif args.play:
+        play_pre_trained(args.parameters_json)

### 2018-01-02 src/agent.py
Fix indentation... :shame:

@@ -222,8 +222,8 @@ class Agent:
 
         # compute epsilon at step t
         completion = self.get_learning_completion()
-		if eps is None:
-        	eps = Parameters.INITIAL_EXPLORATION - (completion * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
+        if eps is None:
+            eps = Parameters.INITIAL_EXPLORATION - (completion * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
         if random.random() < eps:
             # take a random action
             action = np.random.randint(0, self.action_space, size=1)[0]
@@ -254,7 +254,7 @@ class Agent:
 
 
     def play(self):
-		self.environment.render(mode='human')
+        self.environment.render(mode='human')
         while True:
             self.environment.terminal = False
             while self.environment.get_lives():

### 2018-01-02 src/main.py

@@ -48,7 +48,7 @@ if __name__ == "__main__":
     group.add_argument('--plot', action='store_true', help='Plot the results obtained during training')
     group.add_argument('--reset-plot', action='store_true', help='Delete results obtained during training')
     group.add_argument('--random', action='store_true', help='Play 500 games with random action selection and print the mean/std')
-	group.add_argument('--play', action='store_true', help='Play the game with pre-trained model')
+    group.add_argument('--play', action='store_true', help='Play the game with pre-trained model')
 
     args = parser.parse_args()
     assert exists(args.parameters_json)

### 2018-01-02 src/extract_imgs_from_memory.py
Add script to make a mp4 video from memory imgs

@@ -1,0 +1,34 @@
+import numpy as np
+from shutil import copyfile
+from os import remove
+from os.path import exists
+import imageio
+
+MEMORY_PATH = "mem.dat"
+TMP_MEMORY_PATH = MEMORY_PATH + ".copy"
+
+MEM_SIZE = 1000000  #1e6
+SUB_ARRAY_SIZE = 10000  # 1e4
+
+OUTPUT_PATH = 'DQN.mp4'  # TO ADAPT
+
+########## DONT USE
+def copy_file():
+    if exists(TMP_MEMORY_PATH):
+        remove(TMP_MEMORY_PATH)
+    copyfile(MEMORY_PATH, TMP_MEMORY_PATH)
+    MEMORY_PATH = TMP_MEMORY_PATH
+##########
+
+def float16_to_int8(array: np.ndarray):
+    return np.array(np.round(array*255), dtype=np.uint8)
+
+def save_to_file(imgs: int, path: str):
+    output_mp4 = imageio.get_writer(path, mode="I", macro_block_size=42)
+    for img in imgs:
+        output_mp4.append_data(img)
+    output_mp4.close()
+
+memory = np.memmap(MEMORY_PATH, mode="r", shape=(MEM_SIZE, 84, 84), dtype=np.float16)
+memory = float16_to_int8(memory[:SUB_ARRAY_SIZE])
+save_to_file(memory[20:1020], OUTPUT_PATH)

### 2018-01-02 src/agent.py
Make `Agent.get_learning_completion` return 1 if `self.step` is higher than `Parameters.FINAL_EXPLORATION_FRAME`

@@ -209,6 +209,8 @@ class Agent:
         """
         Returns the number of performed learning steps divided by the maximum number of steps
         """
+        if self.step >= Parameters.FINAL_EXPLORATION_FRAME:
+            return 1
         dt_final = Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION
         dt = float(self.step - Parameters.REPLAY_START_SIZE)
         df = float(Parameters.FINAL_EXPLORATION_FRAME - Parameters.REPLAY_START_SIZE)

### 2018-01-02 src/agent.py
Use `random.randint` instead of `np.random.randint` when generating a single number.
```
$ python3.5 -m timeit -s "from numpy.random import randint" "randint(0, 6, size=1)[0]" && python3.5 -m timeit -s "from random import randint" "randint(0, 6)"
1000000 loops, best of 3: 1.94 usec per loop  # np
1000000 loops, best of 3: 1.05 usec per loop  # Py
```

@@ -18,6 +18,11 @@ import sys
 import numpy as np
 import tensorflow as tf
 
+from random import randint  as py_randint  # faster than np for a single element. np is
+                                           # interesting to generate many numbers at once
+
+randint = lambda a, b: py_randint(a, b-1)  # to emulate np randint behaviour: return in [a, b)
+
 class RandomAgent:
     def __init__(self, environment):
         self.action_space = Parameters.GAMES.get_action_space(Parameters.GAME)
@@ -31,7 +36,7 @@ class RandomAgent:
             self.environment.reset()
             self.environment.terminal = False
             while self.environment.get_lives() > 0:
-                action = np.random.randint(0, self.action_space, size=1)[0]
+                action = randint(0, self.action_space)
                 _, reward, done = self.environment.process_step(action)
                 score += reward
                 if done:
@@ -228,7 +233,7 @@ class Agent:
             eps = Parameters.INITIAL_EXPLORATION - (completion * (Parameters.INITIAL_EXPLORATION - Parameters.FINAL_EXPLORATION))
         if random.random() < eps:
             # take a random action
-            action = np.random.randint(0, self.action_space, size=1)[0]
+            action = randint(0, self.action_space)
         else:
             # take a smart action
             input_shape = (1, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)

### 2018-01-02 article/article.tex
[Article] Add 2 plots: average Q-values and score per episode

@@
 
-  \todo{Plots avec les erreurs d'apprentissage (pas de cross-val)}
+\begin{figure*}
+  \includegraphics[width=\textwidth]{figures/dqn_uniform}
+  \caption{(Left) Total score per episode during learning phase, smoothed using a moving average with a window of size 2000.
+     (Right) Estimation of Q, averaged across actions, for each batch during learning phase and smoothed using a moving average
+      with a window of size 2000.}
+\end{figure*}
 

### 2018-01-02 article/article.tex
Cite tensorflow

@@
 Therefore, the NN is trained by minimizing the Huber loss~\citep{huber1964robust} with $\delta=1$ of the Bellman equation (by a -- stochastic -- gradient descent).
-The Huber loss $H_\delta$, parameterized by $\delta \in {\mathbb R^*}^+$, is a continuous function (also $C^1$ but not $C^2$) defined by the MSE
-on a small neighbourhood of $0$, and by the absolute value outside this neighbourhood:
+The Huber loss $H_\delta$, parameterized by $\delta \in {\mathbb R^*}^+$, is a continuous function (also $C^1$ but not $C^2$) defined by the mean
+squared error on a small neighbourhood of $0$, and by the absolute value outside this neighbourhood:
 \begin{equation}
@@
 case of Space Invaders).
+
+\subsubsection{Implementation}
+Tensorflow \todo{explain}
+\citep{tensorflow2015-whitepaper}
 

### 2018-01-02 article/article.tex
[Article] DDQN architecture

@@
 
-\todo{introduce DDDQN}
-
 \section{Methods}
@@
 
-The Neural Network is composed of three CNNs followed by a fully connected layer, leading to the final fully connected layer which is the output.
-Between each of these units is a non-linear unit: a Rectifier Linear Unit (ReLU) \citep{krizhevsky2012imagenet}, but an Exponential Linear Unit (ELU) could
-also be used since it has been shown to be more performant in classification and decision tasks~\citep{DBLP:journals/corr/ClevertUH15}, or even a
-Flexible Rectified Linear Unit~\citep{qiu2017flexible}.
+The Deep Q-Network is a Convolutional Neural Network (CNN) composed of three convolutional layers followed by a fully-connected layer, 
+leading to the final output fully-connected layer.
+Between each of these units a non-linear function is applied, which result in Rectifier Linear Units (ReLU) \citep{krizhevsky2012imagenet}, but Exponential Linear Units (ELU) could
+also be used since it has been shown to be more performant in classification and decision tasks~\citep{DBLP:journals/corr/ClevertUH15}, but also
+Flexible Rectified Linear Units~\citep{qiu2017flexible}.
 
-The first CNN outputs 32 $20 \times 20$ images from a $84 \times 84$ image, using $8 \times 8$ filters and a stride of $4$. The second one outputs
-64 $9 \times 9$ images from the previous 32, using filters of size $4 \times 4$ and a stride of $2$; and the third and last CNN outputs performs
-64 $3 \times 3$ filters with stride $1$ on the 64 subimages of the previous layer to output $7 \times 7$ images.
+The first convolutional layer outputs 32 $20 \times 20$ images from a $84 \times 84$ image, using $8 \times 8$ filters and a stride of $4$. The second one outputs
+64 images of size $9 \times 9$ from the previous 32, using filters of size $4 \times 4$ and a stride of $2$; and the third and last convolutional layer
+applies 64 filters of size $3 \times 3$ with stride $1$ on the 64 subimages of the previous layer to output $7 \times 7$ images.
 
-These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully connected layer with 512 neurons,
-which is then connected to the last layer (again fully connected) which corresponds to the output with the same size as the action space (6 in the
+These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully-connected layer with 512 neurons,
+which is then connected to the last layer (again fully-connected) which corresponds to the output with the same size as the action space (6 in the
 case of Space Invaders).
 
+We also reimplemented a Dueling Deep Q-Network (DDQN) in which the high-level representation of states and action values are separated.
+For this purpose,the same convolutional architecture as before was kept, but the output of the last convolutionel layer was split in two streams of
+equal sizes (each stream input size is the same as the output size of the last convolutional layer), then the resulting tensors have been flattened.
+The first stream, which is designed to represent the potential advantage per action, is fully-connected to 6 neurons (one per action), while the second stream
+contains a dense representation of the state value and is fully-connected to a single neuron.
+
+The last layer of the DDQN outputs Q-value estimates by evaluating the following formula:
+\begin{multline}
+    Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \\ ( A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \Sigma_{a'} A(s, a'; \theta, \alpha) )
+\end{multline}
+where $V$ is the output of the state-value stream, $A$ is the output of the advantage-per-action stream, $s$ is the input state (image), 
+$\alpha$ are the parameters of the state-value stream's fully-connected layer, $\beta$ are the parameters of the state-value stream's
+fully-connected layer, $\theta$ are the parameters of the rest of the network, and $\mathcal{A}$ is the discrete action space.
+
 \subsubsection{Implementation}
-Tensorflow \todo{explain}
-\citep{tensorflow2015-whitepaper}
+The neural networks have been implemented using the software library Tensorflow \citep{tensorflow2015-whitepaper}.
+Thus our models consist of data flow graphs designed in a symbolic programming fashion, where each node of a graph is an operation
+and each edge is a data tensor.
 
-\subsection{Update of the weights}
+\subsection{Weight update}
 
@@
 network is the one used in action prediction, and the second one denotes $\hat Q$ and called the \textit{target action-value} network is used when
-updating the weights. The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of $\hat Q$ are denoted by $\theta^-$.
+updating the weights. This particular method holds for both DQN and DDQN The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of $\hat Q$ are denoted by $\theta^-$.
 
 Every 4 steps, at time $t$, a minibatch of 32 past experiences $\left\{(s_{t_j}, a_{t_j}, r_{t_j}, s_{t_j+1})\right\}_{j=1}^{32}$ is selected uniformly from the
-memory in order to perform a Stochastic Gradient Descent (SGD), meaning that the average of the gradients of the loss function on each experience is used to update
+memory in order to perform a Stochastic Gradient Descent (SGD), thus the average of the gradients of the loss function on each experience is used to update
 the weights. As mentioned above, the loss function is a Huber loss function and is defined as:
@@
   \includegraphics[width=\textwidth]{figures/dqn_uniform}
-  \caption{(Left) Total score per episode during learning phase, smoothed using a moving average with a window of size 2000.
-     (Right) Estimation of Q, averaged across actions, for each batch during learning phase and smoothed using a moving average
+  \caption{(Left) Total score per episode during a learning phase of 1.000.000 steps, smoothed using a moving average with a window of size 2000.
+     (Right) Estimation of Q, averaged across actions, for each batch during a learning phase of 1.000.000 steps and smoothed using a moving average
       with a window of size 2000.}

### 2018-01-02 src/memory.py
Bug fix: random sampling in prioritized memory

@@ -265,7 +265,7 @@ class PrioritizedMemory(Memory):
     def update_probs_and_weights(self):
         probs = self.sampling_probs[:self.memory_usage]
         self.sampling_probs[:self.memory_usage] = probs = probs / probs.sum()
-        max_w_i = self.i_s_weights[:self.memory_usage]
+        max_w_i = self.i_s_weights[:self.memory_usage].max()
         self.i_s_weights[:self.memory_usage] = (self.memory_usage * probs) ** (-self.beta) / max_w_i
 
     def update(self, memory_indices, q_estimates, losses, completion):

### 2018-01-02 src/agent.py
Implement short term memory

@@ -150,8 +150,10 @@ class Agent:
 
                 self.step += 1
 
-                if self.step % 1000 == 0:
+                if self.step % 5000 == 0:
                     self.save_session()
+                if self.step % Parameters.SHORT_TERM_MEMORY_UPDATE_PERIOD == 0:  # This name is waaaaaaaaaaaaaaaay too long <3
+                    self.memory.update_short_term()
 
         self.save_session()
 

### 2018-01-02 src/memory.py

@@ -9,8 +9,12 @@ import os
 from shutil import copyfile, move
 import shelve
 
+from time import time
+
 DEFAULT_MEMMAP_PATH = "mem.dat"
 DEFAULT_SAVE_PATH = "memory-" + Parameters.GAME + '.shelf'
+
+STATE_TYPE = np.float16
 
 class ShortTermMemory:
 
@@ -19,17 +23,92 @@ class ShortTermMemory:
         :param mmap: np.memmap
             Memory map that holds all the experience samples.
             This map represents the long-term memory.
+        :param memory: Memory
+            Parent memory
         """
+        self.history_length = Parameters.AGENT_HISTORY_LENGTH
         self.parent_memory = memory
         self.long_term_mem = mmap
+        self.buff_size = Parameters.SHORT_TERM_MEMORY_SIZE
         buffer_shape = list(mmap.shape)
-        buffer_shape[0] = Parameters.SHORT_TERM_MEMORY_SIZE
-        self.buffer = np.zeros(shape=buffer_shape)
+        buffer_shape[0] = self.buff_size * self.history_length
+        # shape is (history size * STM size [4*50k], img height [84], img width [84]) [~2.7Gb]
+        self.screens_buffer = np.zeros(shape=buffer_shape, dtype=STATE_TYPE)
+        self.use_buffer = False
+        self.minibatch_size = Parameters.MINIBATCH_SIZE
+        self.state_shape = (self.minibatch_size, Parameters.IMAGE_HEIGHT, Parameters.IMAGE_WIDTH, Parameters.AGENT_HISTORY_LENGTH)
+        self.state_t = np.empty(self.state_shape, dtype=STATE_TYPE)
+        self.state_t_plus_1 = np.empty(self.state_shape, dtype=STATE_TYPE)
 
 
     def sample_random(self):
-        random_indices = np.random.choice(self.parent_memory.memory_usage, self.buffer.shape[0])
-        self.buffer = self.long_term_mem[random_indices, :]
+        """
+        Sample a random subset of the memmap to keep in RAM
+        """
+        a = time()
+        self.random_indices = np.random.choice(self.parent_memory.memory_usage - self.history_length, self.buff_size)
+        self.random_indices.sort()
+        buffer_indices = np.arange(1, self.buff_size+1)*self.history_length
+        for offset in range(1, self.history_length+1):
[…]

### 2018-01-03 src/memory.py
Remove stupid removal of dat file
(and never stupidly remove dat ass)

@@ -88,11 +88,6 @@ class Memory:
             self.state_t = np.empty(self.state_shape, dtype=np.float16)
             self.state_t_plus_1 = np.empty(self.state_shape, dtype=np.float16)
             print('Created new memory')
-            try:
-                os.remove(self.memory_filepath)
-                os.remove(self.memory_filepath+'.bak')
-            except FileNotFoundError:
-                pass  # remove files if they exist so if they don't exist, just ignore
             return False
 
 

### 2018-01-03 src/agent.py
Change handling of frames skipping from environment to agent
:cactus: pour Antoine
:1st_place_medal: pour Cédric

<3

@@ -89,7 +89,7 @@ class Agent:
 
     def save_session(self):
         time_at_start_save = time.time()
-        print('{}: [Step {}k  --  Took {:3.2f} s]'.format(datetime.now(), self.step/1000, time_at_start_save - self.last_time), end=' ')
+        print('{}: [Step {}k  --  Took {:3.2f} s]'.format(datetime.now(), self.step//1000, time_at_start_save - self.last_time), end=' ')
         self.last_time = time_at_start_save
 
         Parameters.CURRENT_STEP = self.step
@@ -150,7 +150,7 @@ class Agent:
 
                 self.step += 1
 
-                if self.step % 1000 == 0:
+                if self.step % 5000 == 0:
                     self.save_session()
 
         self.save_session()
@@ -200,7 +200,7 @@ class Agent:
         self.memory.add(screen, action, reward_clipper(reward), terminal,)
 
         # if we started learning
-        if(self.step > Parameters.REPLAY_START_SIZE):
+        if(self.step > Parameters.REPLAY_START_SIZE and self.step % Parameters.FRAME_SKIPPING == 0):
 
             # Perform SGD updates at frequency [Parameters.UPDATE_FREQUENCY]
             if(not(self.step % Parameters.UPDATE_FREQUENCY)):
@@ -226,7 +226,8 @@ class Agent:
         The agent uses its experience and expertise acquired
         through deep learning to make intelligent actions (sometimes)
         """
-
+        if self.step > 0 and self.step % Parameters.FRAME_SKIPPING != 0:
+            return self.last_action
         # compute epsilon at step t
         completion = self.get_learning_completion()
         if eps is None:
@@ -242,7 +243,7 @@ class Agent:
 
             q_values = self.tf_session.run(self.dqn.q_values, {self.dqn_input: dqn_input})
             Plotter.add_q_values_at_t(q_values)
-
+        self.last_action = action
         return(action)
 
 

### 2018-01-03 src/environment.py

@@ -93,31 +93,13 @@ class Environment:
         """
 
         lives_before_action = self.lives
-        cumulated_reward = Parameters.NO_REWARD
-
-        # frame skipping (see Parameters for more information)
-        skipped = 0 # break is evil
-        while(skipped < Parameters.FRAME_SKIPPING and not self.terminal):
-
-            self.take_action(action)
-            cumulated_reward += self.reward
-            self.episode_score += self.reward
-
-            """
-            [Article] For games where there is a life counter, the Atari
-            2600 emulator also sends the number of lives left in the game, which is then used to
-            mark the end of an episode during training.
-            """
-            if self.lives < lives_before_action:
-                cumulated_reward += Parameters.NEGATIVE_REWARD
-                self.terminal = True
-
-                Plotter.add_episode_score(self.episode_score)
-                self.episode_score = 0
-
-            skipped += 1
-
-        self.reward = cumulated_reward
+        self.take_action(action)
+        self.episode_score += self.reward
+        if self.lives < lives_before_action:
+            self.reward += Parameters.NEGATIVE_REWARD
+            self.terminal = True
+            Plotter.add_episode_score(self.episode_score)
+            self.episode_score = 0
 
         return(self.state, self.reward, self.terminal)
 

### 2018-01-03 src/games.py

@@ -13,8 +13,9 @@ class Games:
         self.PONG = "Pong-v0"
         self.SPACE_INVADERS = "SpaceInvaders-v0"
         self.TENNIS = "Tennis-v0"
+        self.BREAKOUT = "Breakout-v0"
 
-        self.available_games_list = [self.ASTEROIDS, self.PONG, self.SPACE_INVADERS, self.TENNIS]
+        self.available_games_list = [self.ASTEROIDS, self.PONG, self.SPACE_INVADERS, self.TENNIS, self.BREAKOUT]
         self.action_space = {}
 
     def define_action_spaces(self):

### 2018-01-07 src/agent.py
Prioritized memory without memory mapped file

@@ -4,7 +4,7 @@
 
 from utils import reward_clipper
 from environment import Environment
-from memory import Memory, PrioritizedMemory, BalancedMemory
+from memory import Memory, PrioritizedMemory
 from parameters import Parameters
 from plot import Plotter
 from dqn import DQN
@@ -113,7 +113,7 @@ class Agent:
                 time_at_start_save - self.last_time))
         self.last_time = time_at_start_save
 
-        Parameters.CURRENT_STEP = self.step
+        Parameters.add_attr("CURRENT_STEP", self.step)
         a = time.time()
         Parameters.update()
         b = time.time()

### 2018-01-07 src/memory.py

@@ -131,29 +131,9 @@ class ShortTermMemory:
         return selected_memories, self.state_t, self.state_t_plus_1
 
 
-class ShortTermBalancedMemory(ShortTermMemory):
-
-    def __init__(self, *args, **kwargs):
-        ShortTermMemory.__init__(self, *args, **kwargs)
-
-    def sample_memory(self, nb_samples=1):
-        """
-        :param nb_samples: int
-            Number of integers to return
-        """
-        rewards = self.rewards[:self.buff_size - 1]
-        weights = np.ones(self.buff_size - 1, dtype=np.float)
-        weights[rewards < 0] *= 15
-        weights[rewards > 0] *= 10
-        weights /= weights.sum()
-        indices = np.random.choice(self.buff_size - 1, nb_samples, p=weights)
-        # print(list(self.rewards[indices]))
-        return indices
-
-
 class Memory:
 
-    def __init__(self, destination=DEFAULT_MEMMAP_PATH, load=True, stm_type=ShortTermMemory):
+    def __init__(self, destination=DEFAULT_MEMMAP_PATH, load=True, use_memmap=True, stm_type=ShortTermMemory):
         """
         :param destination: str
             Path to the file where the long-term experience must be stored
@@ -164,25 +144,31 @@ class Memory:
         """
 
         self.memory_filepath = destination
-
+        self.use_memmap = use_memmap
         self.memory_size = Parameters.LONG_TERM_MEMORY_SIZE
 
         screens_shape = (
             self.memory_size,
             Parameters.IMAGE_HEIGHT,
             Parameters.IMAGE_WIDTH)
-        if os.path.isfile(self.memory_filepath):
-            self.screens = np.memmap(
-                self.memory_filepath,
-                mode="r+",
-                shape=screens_shape,
-                dtype=STATE_TYPE)
+        if self.use_memmap:
+            if os.path.isfile(self.memory_filepath):
+                self.screens = np.memmap(
+                    self.memory_filepath,
+                    mode="r+",
+                    shape=screens_shape,
+                    dtype=STATE_TYPE)
+            else:
+                self.screens = np.memmap(
+                    self.memory_filepath,
+                    mode="w+",
+                    shape=screens_shape,
+                    dtype=STATE_TYPE)
         else:
-            self.screens = np.memmap(
-                self.memory_filepath,
-                mode="w+",
-                shape=screens_shape,
-                dtype=STATE_TYPE)
+            if os.path.isfile(self.memory_filepath):
[…]

### 2018-01-07 article/article.tex
Get further into the article

@@
 \usepackage{alifexi}
+\usepackage{float}
+\usepackage{subcaption}
 \usepackage{color}
@@
 Universit\'{e} Libre de Bruxelles \\
-\{Cedric.Simar, apassemi, sgueniff, robpetit\}@ulb.ac.be}
+\{csimar, apassemi, sgueniff, robpetit\}@ulb.ac.be}
 
@@
 The huge breakout effect of the DQN was that it has shown to be a first step towards general artificial intelligence~\citep{togelius2015ai}, meaning an
-algorithm that is not only good at learning one particular task but several of them. Even though the DQN needs to be trained for about a million steps
-on each Atari game, it gets good at most of them and better than human at 22 of the 49 experimented~\citep{mnih2015human}.
+algorithm that is not only good at learning one particular task but several of them. Even though the DQN needs to be trained for tens of millions steps
+on each Atari game, it gets good at most of them and better than human (considered $75\%$ of human score or more) at 22 of the 49 experimented~\citep{mnih2015human}.
+
+Finally it has been pointed out by the original DQN conceptors that DQN has an ability to learn by acquiring experience generated from policies
+other than its own. \todo{On peut en faire qqch pas vrai ?}
+
+\section{Methods}
+
+Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}.
+
+We developed our models with the support of the Arcade Learning Environment~\citep{bellemare13arcade}. This gives the opportunity to either
+retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
+applications). Because of the very high dimensionality of raw images, we applied some preprocessing steps on them. These steps are also aimed at
+removing some artifacts produced by the Atari architecture. More specifically, the flickering effect implies that some game sprites only appear
+in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
+Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance, using the following formula:
+\begin{equation}
+    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B.
+\end{equation}
+
+After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
+$84 \times 84$ pixel grid.
 
@@
 \end{equation}
-where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
-Space Invaders professional gamer \textbf{Stanislas Gueniffey}. By normalizing in this manner, the random action selection score RAS results in
-a normalized performance of 0\% and the human score results in a normalized performance of 100\%.
-
-Finally it has been pointed out by the original DQN conceptors that DQN has an ability to learn by acquiring experience generated from policies
-other than its own. \todo{On peut en faire qqch pas vrai ?}
-
-\section{Methods}
-
-Our source code is available at \url{https://github.com/RobinPetit/Learning-Chaos}.
-
-We developed our models with the support of the Arcade Learning Environment~\citep{bellemare13arcade}. This gives the opportunity to either
-retrieve the environment state as a raw image (for deep reinforcement learning research) or as a state of the RAM (for classic reinforcement learning
-applications). Because of the very high dimensionality of raw images, we applied some preprocessing steps on them. These steps are also aimed at
-removing some artifacts produced by the Atari architecture. More specifically, the flickering effect implies that some game sprites only appear
-in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
-Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance, using the following formula:
-\begin{equation}
-    \text{Luminance} = 0.299 \times R + 0.587 \times G + 0.114 \times B.
-\end{equation}
-
-After mapping the three color channels to a unique channel, the number of pixels has also been reduced. Each input image as been rescaled to a
-$84 \times 84$ pixel grid.
+where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by Google's professional
+game tester after a training period of 2 hours.  By normalizing in this manner, the random action selection score RAS results in a normalized performance
+of 0\% and the human score results in a normalized performance of 100\%.
 
@@
 
-The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and has (\todo {probably}) been shown to increase
-stability.~\todo{check if this has a proper reference or drop it}
+The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and is claimed to increase stability~\citep{mnih2015human},
+yet no further investigation has been performed.
 
@@
 These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully-connected layer with 512 neurons,
-which is then connected to the last layer (again fully-connected) which corresponds to the output with the same size as the action space (6 in the
-case of Space Invaders).
+which is then connected to the last layer (again fully-connected) which corresponds to the output with the same size as the action space (4 in the
+case of Breakout: \textit{NOOP}, \textit{FIRE}, \textit{LEFT} and \textit{RIGHT}).
 
@@
 \begin{multline}
-    Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \\ ( A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \Sigma_{a'} A(s, a'; \theta, \alpha) )
+    Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \\ \left(A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \sum_{a'} A(s, a'; \theta, \alpha)\right)
 \end{multline}
@@
 network is the one used in action prediction, and the second one denotes $\hat Q$ and called the \textit{target action-value} network is used when
-updating the weights. This particular method holds for both DQN and DDQN The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of $\hat Q$ are denoted by $\theta^-$.
+updating the weights. This particular method holds for both DQN and DDQN The parameters of $Q$ are denoted by $\theta$ as previously, and the parameters of
+$\hat Q$ are denoted by $\theta^-$.
 
@@
 \end{equation}
-with $y_j$ set either to $r_j$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
-$r_j + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.
+with $y_j$ set either to $r_{t_j}$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
+$r_{t_j} + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{t_j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.
 
@@
 
-\begin{figure*}
-  \includegraphics[width=\textwidth]{figures/dqn_uniform}
-  \caption{(Left) Total score per episode during a learning phase of 1.000.000 steps, smoothed using a moving average with a window of size 2000.
-     (Right) Estimation of Q, averaged across actions, for each batch during a learning phase of 1.000.000 steps and smoothed using a moving average
-      with a window of size 2000.}
+\begin{figure*}[!t]
+	\vspace{-1cm}
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/dqn_uniform_e_scores}
+		\subcaption{Total score per episode during a learning phase of 16,000,000 steps, smoothed using a moving average with a window of size 1000.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/dqn_uniform_q_values}
+		\subcaption{Estimation of $Q$, averaged across actions, for each batch during a learning phase of 16,000,000 steps and smoothed using a moving
+			average with a window of size 5000.}
+	\end{subfigure}
+	\caption{Episode score and $Q$ values estimation of the DQN after 16,000,000 steps of training on Breakout.\label{fig:dqn uniform}}
 \end{figure*}
 
-Figure~\ref{fig:t-SNE}\todo{Link the right figure} shows a 2-dimensional representation of the last fully connected hidden layer, i.e. a 512-dimensional
+Figure~\ref{fig:dqn uniform} shows the results of the DQN after 16,000,000 steps of training. One can observe that the first 15,000 batches didn't help
+increasing the $Q$ estimates: the internal representation was being built up. Also, the score per episode is very low at first since the agent plays
+randomly, a million steps were required for $\epsilon$ to linearly decrease from $1$ to $0.1$.
+
+After this phase, the estimates increase and so does the score per episode, until they both reached stability around 11,000,000 steps.
+
+\begin{table}[!h]
+	\centering
+	% std of Random is 1.27 & std of DQN is 22.81
+	\begin{tabular}{l|l|l|l}
+		Model & Random & Human & DQN \\ \hline \hline
+		Score & $1.31$ & $31.8$ & $39.88$ \\ \hline
+		P &  $0\%$ & $100\%$ & $126.83\%$
+	\end{tabular}
+	\caption{Score and Performance on Breakout (averaged on 100 simulations with $\varepsilon=5\%$ and with one action decision per frame). The DQN has been
+	trained during 16,000,000 steps.
+	\todo{Ajouter scores Prioritized + DDQN et le temps d'apprentissage}\label{tab:scores breakout}}
+\end{table}
+
+Table~\ref{tab:scores breakout} shows that the DQN performed better than human after 16,000,000 training steps, with a performance of $126.83\%$.
+
+Figure~\ref{fig:t-SNE DQN uniform} shows a 2-dimensional representation of the last fully connected hidden layer, i.e. a 512-dimensional
 vector. The method of \textit{projection} that was used is t-SNE~\citep{maaten2008visualizing}, which is part of the manifold learning algorithms, also
@@
 
-  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop}~\citep{wattenberg2016how}
+  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop. Robin est dessus.}~\citep{wattenberg2016how}
+
+\begin{figure}
+	\includegraphics[width=.5\textwidth]{figures/TSNE_dqn_uniform}
+	\caption{\todo{Replace with correct plot}.\label{fig:t-SNE DQN uniform}}
+\end{figure}
+
+Figure~\ref{fig:dqn uniform scores dist} shows the distribution of scores when letting the DQN play for 100 games. We can see that despite having an
+average score of \todo{XXX}, the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered enough for the
+DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
+and more importantly, the agent selects the action at each frame. This highest score of 219 is probably due to a non-anticipated breaking through
+the brick layer, leading to the ball destroying the upper layer.
+
+\begin{figure}
+	\includegraphics[width=.5\textwidth]{figures/dqn_uniform_score_distribution}
+	\caption{Distribution of the scores performed by the DQN on 100 plays.\label{fig:dqn uniform scores dist}}
+\end{figure}
+
+\begin{figure*}{!b}
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/dqn_uniform_e_scores}
+		\subcaption{Total score per episode of the DDQN during its learning phase, smoothed using a moving average of size \todo{XXX}.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/dqn_uniform_q_values}
+		\caption{Estimation of $Q$ by the DDQN, averaged across actions, for each batch during its learning phase and smoothed using a moving average
+		with a window of size \todo{XXX}.}
+	\end{subfigure}
+	\caption{\todo{Ajouter les plots DDQN et remplacer les paths des includegraphics}.\label{fig:ddqn uniform}}
+\end{figure*}
 
@@
   \todo{Amélioration: Double dueling DQN}~\citep{DBLP:journals/corr/WangFL15}
+
+\section{Conclusion}
+
+	\todo{Ajouter ordre de performance des 3 tests}
 

### 2018-01-08 src/plot.py
Make t-SNE plots bloody great again

@@ -6,6 +6,8 @@ import os
 import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
+from matplotlib.patches import ConnectionPatch
+import cv2
 
 import gym
 from parameters import Parameters
@@ -126,6 +128,7 @@ class Plotter:
         folder = "./out/layers_plots/" + str(Parameters.GAME) + "/" + str(agent.step) + "/"
 
         filters = {"conv1": 32, "conv2": 64, "conv3": 64}
+        filters_scale = {"conv1": 10, "conv2": 20, "conv3": 30}
 
         if not os.path.exists(folder):
             os.makedirs(folder)
@@ -148,8 +151,10 @@ class Plotter:
 
             layer_output = agent.tf_session.run(agent.dqn.layers[layer], {agent.dqn_input: state_t})
 
+            scale = filters_scale[layer]
             for filter in range(filters[layer]):
-                plt.imsave(layer_folder + layer + "_filter_" + str(filter+1) + ".png", layer_output[img,:,:,filter], cmap='gray')
+                plt.imsave(layer_folder + layer + "_filter_" + str(filter+1) + ".png",
+                    cv2.resize(layer_output[img,:,:,filter], (0,0), fx=scale, fy=scale, interpolation=cv2.INTER_AREA), cmap='gray')
 
 
 
@@ -157,12 +162,46 @@ class EmbeddingProjector(TSNE):
     def __init__(self, *args, **kwargs):
         TSNE.__init__(self, *args, **kwargs)
 
-    def save_plot(self, hidden_repr, v_values, folder):
+    def save_plot(self, states, hidden_repr, v_values, folder):
         projected = self.fit_transform(hidden_repr)
-        degrees = (v_values - v_values.min()) / \
-            (v_values.max() - v_values.min())
-        colors = matplotlib.cm.jet(degrees)
-        cs = [colors[i] for i in range(len(projected))]
-        plt.scatter(projected[:, 0], projected[:, 1], color=cs)
-        plt.title("t-distributed Stochastic Neighbor Embedding")
-        plt.savefig(os.path.join(folder, "tsne.png"))
+        print('projected\n\tCreating display')
+
+        # create different axes of different size
+        ax_tsne = plt.subplot2grid((2, 3), (0, 0), rowspan=2, colspan=2)
[…]

### 2018-01-08 article/article.tex
[Article] <3

@@
 
-\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[draft,colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
 \usepackage{flushend}
+
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
 
@@
 
+One of the big advantages of deep reinforcement learning is that it relies on a general-purpose learning procedure which allows to make
+abstraction of each game's specificities. Although this project focuses on performing on Breakout (a popular Atari game), we made sure
+that our agents could be used to learn playing on any Atari game.
+
 \section{Methods}
@@
 $84 \times 84$ pixel grid.
-
-Since we are handling visual features, we used screen frames as environment state representations. Each frame is a (210 x 160) color image sampled at
-a frequency of 60 Hz. We denote the environment state at current step $t$ by $s_t$. The agent gathers experience by incorporating new experience samples
-to its history. An experience is a tuple $e_t = (s_t, a_t, r_t, s_{t+1})$ appended to the experience history immediately after being generated.
-To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse to the experience
-replay algorithm ~\citep{adam2012experience}. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
-We compared this technique to a more efficient approach called prioritized experience replay~\citep{DBLP:journals/corr/SchaulQAS15}.
-To emphasize the importance of the memory replay mechanism, we compared the results obtained using prioritized experience replay, regular
-experience replay, and no experience replay.
 
@@
 
-We also reimplemented a Dueling Deep Q-Network (DDQN) in which the high-level representation of states and action values are separated.
+In addition to the original DQN article, we also reimplemented a Dueling Deep Q-Network \citep{DBLP} (DDQN) in which the high-level representation of states 
+and action values are separated. This shouldn't be confused with the notion of target neural network. Indeed each of our model is copied to a target model
+of the same architecture. The DDQN accompanied by its target DDQN, constitute a whole different agent.
 For this purpose,the same convolutional architecture as before was kept, but the output of the last convolutionel layer was split in two streams of
@@
 \subsubsection{Implementation}
-The neural networks have been implemented using the software library Tensorflow \citep{tensorflow2015-whitepaper}.
+All CNNs have been implemented using the software library Tensorflow \citep{tensorflow2015-whitepaper}.
 Thus our models consist of data flow graphs designed in a symbolic programming fashion, where each node of a graph is an operation
@@
 The $\hat Q$ target action-value network is then copied back from the action value network $Q$ every 10,000 steps.
+
+\subsection{Hierarchical memory}
+
+Since we are handling visual features, we used screen frames as environment state representations. Each frame is a (210 x 160) color image sampled at
+a frequency of 60 Hz. We denote the environment state at current step $t$ by $s_t$. The agent gathers experience by incorporating new experience samples
+to its history. An experience is a tuple $e_t = (s_t, a_t, r_t, s_{t+1})$ appended to the experience history immediately after being generated.
+To prevent the agent from forgetting the history and increase the bias towards most recent experience samples, we have recourse to the experience
+replay algorithm ~\citep{adam2012experience}. One can straightforwardly implement it as a uniform sampling algorithm that returns random experience samples at each step.
+We compared this technique to a more efficient approach called prioritized experience replay~\citep{DBLP:journals/corr/SchaulQAS15}.
+To emphasize the importance of the memory replay mechanism and as group contribution, we compared the results obtained using prioritized experience replay, uniform
+experience replay, and no experience replay.
+
+Regular experience replay consists in feeding the CNN with uniformly random samples stored in memory. To ensure the agent remembers experience samples $e_t$ that
+occured long time ago ($t$ being significantly less than current step), one must allow the replay memory to be really large. Most researchers use a memory size
+of 1,000,000 samples, which is too large for most personnal computers (over 16 Gb in the framework of our project). Our second and main contribution consisted in
+overcoming this issue by implementing a hierarchical style of memory and showing that it does not jeopardize the performance of Deep Q-networks.
+
+Our hierarchical memory consists of a short-term memory and a long-term memory. The latter stores all the experience samples but is not directly accessed by the agent.
+Every time the agent needs to perform a mini-batch gradient descent, it requests a mini-batch to the short-term memory. To ensure diversity of the experience,
+the short-term memory is updated by copying contiguous data from the long-term memory, where the beginning of the contiguous selected area is selected randomly.
+This update operation is repeated every 30,000 steps only in order to avoid a significant reduction in learning speed.
+
+\subsubsection{Implementation}
+
+Long-term experience samples are stored in a large-sized memory-mapped file. The corresponding memory map is used as an abstract data structure that allows to access
+the samples by indexing. Short-term memory is implemented as a contiguous array. This array is regularly updated by copying contiguous chunks of data samples from
+long-term memory. The agent receives random mini-batches constructed from this short-term array. A contrario, each new experience sample acquired by the agent
+(via the environment) is stored directly into the long-term memory. Indeed it appeared that the time required to add a new sample to memory was negligible 
+compared with the time required to generate a mini-batch of 32 samples. Actually because the random selection is uniform (for regular experience replay),
+the selected samples have very large gaps between them, which causes a lot of cache misses.
+
+\subsubsection{Prioritized experience replay}
+
+Prioritized experience replay~\citep{DBLP:journals/corr/SchaulQAS15} has been shown to drastically reduce learning time requirements for many deep
+reinforcement learning problems, or even improve model performance in some problems, in particular a large majority of Atari games.
+The main shortcoming of uniform experience replay is the fact that mini-batches may contain highly redundant states. Plus, most experience samples have
+a zero reward. Gradient descent does not perform well when only negative samples are used. We used rank-based prioritization, which consists in ranking
+experience samples by how well they "surprised" the model, and randomly selecting them with respect to weights that are positively correlated with the ranks.
+
+Schaul et al. introduced different ways to compute these ranks. We chose to prioritize with the temporal-difference error:
+\begin{multline}
+    \delta_t = r_t + \gamma_t Q_{target}(s_t, \argmin_a Q(s_t, a)) - Q(s_{t-1}, a_{t-1})
+\end{multline}
+where $Q_{target}$ is a function that applies a forward pass using the weights of the target CNN, $Q$ is a similar function
+but for the first CNN, and $r_t$ is the reward received at step $t$.
+
+To create a mini-batch, samples are selected according to a weighted distribution, where the weights are computed as follows:
+\begin{multline}
+    P(j) = p_i^{alpha} / \Sigma_i p_i^{alpha}
+\end{multline}
+$p_i$ designates the absolute value of the temporal-difference error on particular sample $i$ at current step $t$.
+At the start of learning phase the vast majority of probabilities $p_i$ haven't been evaluated yet. This is why we initialized
+them with a value of $1$ (which corresponds to a uniform random selection). Coefficient $\alpha$ is a hyper-parameter that
+determines how strong the prioritization is. If $\alpha$ is set to zero, one falls back on the case of uniform random selection.
+
+To take into account the relative importance of each sample inside a mini-batch, we had to modify the DQN learning procedure and weight the Huber loss
+by an importance sampling weight, calculated from the probabilities $p_i$. For more details about importance sampling weights, please refer to the original
+article. The difference between our implementation and what was described in the article is the utilization of the Huber loss instead of only
+the temporal-difference error for updating weights.
 
@@
 
-  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop. Robin est dessus.}~\citep{wattenberg2016how}
+  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop. Robin est dessus. Robin gère.}~\citep{wattenberg2016how}
 
@@
 
-  \todo{Amélioration: Double dueling DQN}~\citep{DBLP:journals/corr/WangFL15}
+\todo{Les CNNs c'est instable, faut du temps et faire du café et bcp de parameter tuning -> solution plus stable ? Modèle linéaire (cf article d'origine) ?}
+
+\todo{Comme pour la prioritized faut mettre à jour les poids de tous les samples à chaque step, c'est compliqué faut optimiser ça}
+
+\todo{LSTM ?}
 

### 2018-01-08 article/article.tex
[Article] add plots in results + start analysis

@@
 \usepackage{float}
+\usepackage[font=small,labelfont=bf]{caption}
 \usepackage{subcaption}
@@
 
-\usepackage[draft,colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
 \usepackage{flushend}
@@
 \begin{multline}
-    \delta_t = r_t + \gamma_t Q_{target}(s_t, \argmin_a Q(s_t, a)) - Q(s_{t-1}, a_{t-1})
+    \delta_t = r_t + \gamma_t Q_{\text{target}}\big(s_t, \argmin_a Q(s_t, a)\big) - Q(s_{t-1}, a_{t-1})
 \end{multline}
-where $Q_{target}$ is a function that applies a forward pass using the weights of the target CNN, $Q$ is a similar function
+where $Q_{\text{target}}$ is a function that applies a forward pass using the weights of the target CNN, $Q$ is a similar function
 but for the first CNN, and $r_t$ is the reward received at step $t$.
@@
 To create a mini-batch, samples are selected according to a weighted distribution, where the weights are computed as follows:
-\begin{multline}
-    P(j) = p_i^{alpha} / \Sigma_i p_i^{alpha}
-\end{multline}
+\begin{equation}
+    P(j) = p_j^{\alpha} / \sum_i p_i^{\alpha}
+\end{equation}
 $p_i$ designates the absolute value of the temporal-difference error on particular sample $i$ at current step $t$.
@@
 \begin{figure*}[!t]
-	\vspace{-1cm}
+	\vspace{-1.5cm}
 	\begin{subfigure}{.47\textwidth}
@@
 	\centering
-	% std of Random is 1.27 & std of DQN is 22.81
-	\begin{tabular}{l|l|l|l}
-		Model & Random & Human & DQN \\ \hline \hline
-		Score & $1.31$ & $31.8$ & $39.88$ \\ \hline
-		P &  $0\%$ & $100\%$ & $126.83\%$
+	\small
+	\setlength\tabcolsep{5pt}
+	\begin{tabular}{l|l|l|l|l|l}
+		Model & Random & Human & DQN (U) & DQN (P) & DDQN \\ \hline \hline
+		Score & $1.31$ & $31.8$ & $39.88$ & $0.22$ & $30.26$ \\ \hline
+		P &  $0\%$ & $100\%$ & $126.83\%$ & $-3.57\%$ & $94.95\%$
 	\end{tabular}
-	\caption{Score and Performance on Breakout (averaged on 100 simulations with $\varepsilon=5\%$ and with one action decision per frame). The DQN has been
-	trained during 16,000,000 steps.
-	\todo{Ajouter scores Prioritized + DDQN et le temps d'apprentissage}\label{tab:scores breakout}}
+	\caption{Score and Performance on Breakout (averaged on 100 simulations with $\varepsilon=5\%$ and with one action decision per frame). DQN (U) represents the
+	DQN with uniform replay memory, and DQN (P) represents the DQN with prioritized memory. The DQNs have been trained during 16,000,000 steps, and the DDQN has
+	been trained for 7,000,000 steps.\label{tab:scores breakout}}
 \end{table}
@@
 
-The implementation of t-SNE that was used is the one from Scikit-Learn~\citep{scikit-learn}, the Python library.
+The implementation of t-SNE that was used is the one from Scikit-Learn~\citep{scikit-learn}, the Python library. The parameters are the default ones:
+a perplexity of 30 and a learning rate of 200 on 1000 iterations. The learning rate is quite high relatively to the propositions in~\cite{wattenberg2016how},
+which allows to decrease embedding time. The perplexity lies in the range from 5 to 50 which is advised.
 
-  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop. Robin est dessus. Robin gère.}~\citep{wattenberg2016how}
+\todo{Parler du clustering.}
 
-\begin{figure}
+The two windows on the right of the figure show the states leading to the smallest estimated $Q$ value (blue) and the highest (red).
+We can observe that the image with the smallest estimated $Q$ is an image without ball, meaning that no immediate reward is available. It
+is worth noting that the DQN doesn't seem to have learned to use the FIRE action when it misses the ball. The image with the highest estimated
+$Q$ is an image where the ball is about to hit a brick, leading to an immediate positive reward.
+
+\begin{figure}[!h]
 	\includegraphics[width=.5\textwidth]{figures/TSNE_dqn_uniform}
-	\caption{\todo{Replace with correct plot}.\label{fig:t-SNE DQN uniform}}
+	\caption{Projection of 320,000 states onto a 2-dimensional space. The color of each dot (from dark green to yellow) represents the estimated $Q$ value of the
+	best action the agent can play. The states leading to the smallest and the highest estimated $Q$ are displayed with respectively a blue and a red border.
+	\label{fig:t-SNE DQN uniform}}
 \end{figure}
@@
 Figure~\ref{fig:dqn uniform scores dist} shows the distribution of scores when letting the DQN play for 100 games. We can see that despite having an
-average score of \todo{XXX}, the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered enough for the
-DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
+average score of 39.88 (with a standard deviation of 41.14), the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered
+enough for the DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
 and more importantly, the agent selects the action at each frame. This highest score of 219 is probably due to a non-anticipated breaking through
-the brick layer, leading to the ball destroying the upper layer.
+the brick layer, leading to the ball destroying the upper layer whose bricks add 7 points to the score.
 
-\begin{figure}
+\begin{figure}[!h]
 	\includegraphics[width=.5\textwidth]{figures/dqn_uniform_score_distribution}
@@
 
-\begin{figure*}{!b}
+Figure~\ref{fig:DDQN + DQNP} shows the evolution of reward per episode and of $Q$ estimates on 16,000,000 steps for the DDQN and 7,000,000 for the DQN with
+prioritized memory.
+
+One can observe that the DDQN seems to stabilize around an average reward per episode of 1.4 (though more training time would be needed to ensure convergence),
+like the DQN with uniform memory. Yet, $Q$ estimates of the DQN end up around $2$ whereas they are only half as big for the DDQN.
+
+\begin{figure*}[!t]
+	\vspace{-1.5cm}
 	\begin{subfigure}{.47\textwidth}
-		\includegraphics[width=\textwidth]{figures/dqn_uniform_e_scores}
-		\subcaption{Total score per episode of the DDQN during its learning phase, smoothed using a moving average of size \todo{XXX}.}
+		\includegraphics[width=\textwidth]{figures/ddqn_uniform_e_scores}
+		\subcaption{Total score per episode of the DDQN using uniform memory during its learning phase, smoothed using a moving average of size 1000.}
 	\end{subfigure}
@@
 	\begin{subfigure}{.47\textwidth}
-		\includegraphics[width=\textwidth]{figures/dqn_uniform_q_values}
-		\caption{Estimation of $Q$ by the DDQN, averaged across actions, for each batch during its learning phase and smoothed using a moving average
-		with a window of size \todo{XXX}.}
+		\includegraphics[width=\textwidth]{figures/ddqn_uniform_q_values}
+		\caption{Estimation of $Q$ by the DDQN using uniform memory, averaged across actions, for each batch during its learning phase and smoothed using a moving average
+		with a window of size 5000.}
 	\end{subfigure}
-	\caption{\todo{Ajouter les plots DDQN et remplacer les paths des includegraphics}.\label{fig:ddqn uniform}}
+	\begin{subfigure}{.47\textwidth}
+		\vspace{-.15cm}
+		\includegraphics[width=\textwidth]{figures/dqn_prioritized_e_scores}
+		\subcaption{Total score per episode of the DQN with prioritized memory during its learning phase, smoothed using a moving average of size 1000.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{.47\textwidth}
+		\vspace{-.15cm}
+		\includegraphics[width=\textwidth]{figures/dqn_prioritized_q_values}
+		\caption{Estimation of $Q$ by the DQN with prioritized memory, averaged across actions, for each batch during its learning phase and smoothed using a moving average
+		with a window of size 5000.}
+	\end{subfigure}
+	\caption{Adaptation of Figure~\ref{fig:dqn uniform} for both DDQN after 16,000,000 training steps (a and b) and DQN with prioritized memory after 7,000,000
+	training steps (c and d).\label{fig:DDQN + DQNP}}
 \end{figure*}
+
+\begin{figure*}[!t]
+	\vspace{-.3cm}
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/ddqn_uniform_score_distribution}
+		\caption{Distribution of the score performed by the DDQN with uniform memory on 100 plays.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{.47\textwidth}
+		\includegraphics[width=\textwidth]{figures/dqn_prioritized_score_distribution}
+		\caption{Distribution of the score performed by the DQN with prioritized memory on 100 plays.}
+	\end{subfigure}
+	\caption{Adaptation of Figure~\ref{fig:dqn uniform scores dist} for both DDQN (a) and DQN with prioritized memory (b).\label{fig:scores DDQN + DQNP}}
+\end{figure*}
+
+\todo{Continuer (ici pour ne pas flinguer la mise en page) l'analyse comparative}
 
@@
 
-\newpage
+%\newpage
 \footnotesize

### 2018-01-08 article/article.tex
[Article] Finish the article

@@
 \usepackage{float}
-\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{subcaption}
+\usepackage{dblfloatfix}
+\usepackage[font=footnotesize,labelfont=bf]{caption}
+\usepackage[font=scriptsize]{subcaption}
 \usepackage{color}
@@
 
-\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=blue]{hyperref}
 \usepackage{flushend}
@@
 
-  Reinforcement learning is a subarea of machine learning inspired by neuroscience, where learners have to select actions with
+  Reinforcement learning is a subarea of machine learning inspired by neuroscience where learners have to select actions with
   regards to the environment state and receive a reward accordingly. Each agent's objective is to maximize its cumulative reward across its whole lifetime,
@@
   researchers usually deal with a small discrete space to represent the values of Q. This is no longer sufficient if one desires to describe the environment state
-  as an image of raw pixels. In consequence, we introduce deep Q-learning methods, which have been found to be really effective in mapping raw pixels
+  as an image of raw pixels. Hence, we introduce deep Q-learning methods, which have been found to be really effective in mapping raw pixels
   to abstract high-level features. In deep Q-learning, the estimation of Q-values for a given action are produced using deep neural networks,
-  which consist of many neural layers stacked on top of each other.
-  These techniques have revealed themselves to be able to beat human experts at playing Atari games
+  which consists in many neural layers stacked on top of each other.
+  These techniques have proved to be able to beat human experts at playing Atari games
   by learning only from visual features. Despite the current progress, learning how to take complex decisions on the basis of high-dimensional visual data remains an
@@
 A DQN is a particular instance of Convolutional Neural Networks (CNN) and is composed of a stack of neural layers~\citep{lecun1998gradient}. One characteristic of CNNs is the
-presence of convolutional filters that maps pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
-allows the convolutional layer to capture the local information of the pixels, contrary to what dense (fully-connected) layers do. This procedure is greatly
+presence of convolutional filters that map pixel luminance to more abstract features. Each filter (or kernel) is locally connected to its output, which
+allows the convolutional layer to capture the local information of the pixels, as opposed to what dense (fully-connected) layers do. This procedure is greatly
 inspired by the notion of receptive field introduced by Hubel and Wiesel~\citep{Hubel1962}.
@@
 Finally it has been pointed out by the original DQN conceptors that DQN has an ability to learn by acquiring experience generated from policies
-other than its own. \todo{On peut en faire qqch pas vrai ?}
+other than its own.
 
 One of the big advantages of deep reinforcement learning is that it relies on a general-purpose learning procedure which allows to make
-abstraction of each game's specificities. Although this project focuses on performing on Breakout (a popular Atari game), we made sure
-that our agents could be used to learn playing on any Atari game.
+abstraction of each game's specificities. Although this project focuses on performing on Breakout (a popular Atari game),
+our agents could be used to learn playing on most Atari game.
 
@@
 in one out of two frames. An easy solution consists in taking the pixelwise maximum value between the current frame and the previous frame.
-Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance, using the following formula:
+Then the sample size has been reduced by a factor of three by mapping RGB values to the pixel luminance using the following formula:
 \begin{equation}
@@
 
-Every 4 steps of the Q-learning, the agent builds a minibatch from its memory (either one of the three mechanisms we just mentioned):
+Every 16 steps of the Q-learning, the agent builds a minibatch from its memory (either one of the three mechanisms we just mentioned):
 this minibatch is used by the DQN to process one step of stochastic gradient descent. In this way the values of Q are estimated iteratively and
@@
 
-The perceived rewards of the agent have all been reduced to be either $-1$ for a negative reward, $0$ for a session with no reward, and $+1$
+The perceived rewards of the agent have all been clipped to be either $-1$ for a negative reward, $0$ for a session with no reward, and $+1$
 for a positive reward, i.e. at any time $t$: $r_t \in \{-1, 0, +1\}$. No intermediate or larger reward is available to the learning algorithm.
-This allows to learn without any previous knowledge on the particular game the DQN is trained on. Yet some rewards can be highly different (in Pac-Man likes,
+This allows to learn without any previous knowledge of the particular game the DQN is trained on. Yet some rewards can be highly different (in Pac-Man likes,
 eating a cherry gives 100 points, a banana gives 5000, and eating a ghost gives $2^n \times 100$ with $n$ the number of eaten ghosts in a single run).
@@
 
-The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and is claimed to increase stability~\citep{mnih2015human},
+The interest of the Huber loss is to bound the gradient of the loss between $-1$ and $+1$, and is claimed to increase stability~\citep{mnih2015human},
 yet no further investigation has been performed.
@@
 
-In addition to the original DQN article, we also reimplemented a Dueling Deep Q-Network \citep{DBLP} (DDQN) in which the high-level representation of states
+In addition to the original DQN article, we also reimplemented a Dueling Deep Q-Network \citep{DBLP:journals/corr/WangFL15} (DDQN) in which the high-level
+representation of states
 and action values are separated. This shouldn't be confused with the notion of target neural network. Indeed each of our model is copied to a target model
 of the same architecture. The DDQN accompanied by its target DDQN, constitute a whole different agent.
-For this purpose,the same convolutional architecture as before was kept, but the output of the last convolutionel layer was split in two streams of
+For this purpose,the same convolutional architecture as before was kept, but the output of the last convolutional layer was split in two streams of
 equal sizes (each stream input size is the same as the output size of the last convolutional layer), then the resulting tensors have been flattened.
@@
 Regular experience replay consists in feeding the CNN with uniformly random samples stored in memory. To ensure the agent remembers experience samples $e_t$ that
-occured long time ago ($t$ being significantly less than current step), one must allow the replay memory to be really large. Most researchers use a memory size
-of 1,000,000 samples, which is too large for most personnal computers (over 16 Gb in the framework of our project). Our second and main contribution consisted in
+occurred long time ago ($t$ being significantly less than current step), one must allow the replay memory to be really large. Most researchers use a memory size
+of 1,000,000 samples, which is too large for most personal computers (over 7 Gb in the framework of our project). Our second and main contribution consisted in
 overcoming this issue by implementing a hierarchical style of memory and showing that it does not jeopardize the performance of Deep Q-networks.
@@
 
-After this phase, the estimates increase and so does the score per episode, until they both reached stability around 11,000,000 steps.
+After this phase, the estimates increase and so does the score per episode, until they both started to settle around 11,000,000 steps. Even though
+the score per episode seems to slowly stabilize, the model would need more training time to fully reach stability.
+
+As the $Q$ values computed at each batch are the $Q$ values for the images that the DQN observed at the update step, it is expected to not see
+them fully converge: a SGD is performed every 16 steps, independently of the current image. Therefore if the weights update is performed after
+a frame where the agent receives positive reward no matter its action (e.g. right before the ball hits a brick) or after a frame where the agent
+receives negative reward, no matter his action (e.g. if the ball is about to fall and the paddle is too far away), the $Q$ values will be different.
 
@@
 
-\todo{Parler du clustering.}
-
-The two windows on the right of the figure show the states leading to the smallest estimated $Q$ value (blue) and the highest (red).
-We can observe that the image with the smallest estimated $Q$ is an image without ball, meaning that no immediate reward is available. It
-is worth noting that the DQN doesn't seem to have learned to use the FIRE action when it misses the ball. The image with the highest estimated
-$Q$ is an image where the ball is about to hit a brick, leading to an immediate positive reward.
-
 \begin{figure}[!h]
@@
 
-Figure~\ref{fig:dqn uniform scores dist} shows the distribution of scores when letting the DQN play for 100 games. We can see that despite having an
-average score of 39.88 (with a standard deviation of 41.14), the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered
-enough for the DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
-and more importantly, the agent selects the action at each frame. This highest score of 219 is probably due to a non-anticipated breaking through
-the brick layer, leading to the ball destroying the upper layer whose bricks add 7 points to the score.
+The states end up clustered by images with similar $Q$ values estimates, so the DQN estimates close $Q$ values for similar images.
+
+The two windows on the right of the figure show the states leading to the smallest estimated $Q$ value (blue) and the highest (red).
+We can observe that the image with the smallest estimated $Q$ is an image without ball, meaning that no immediate reward is available. It
+is worth noting that the DQN doesn't seem to have learned to use the FIRE action when it misses the ball. The image with the highest estimated
+$Q$ is an image where the ball is about to hit a brick, leading to an immediate positive reward.
+
+It is interesting to notice that the image with the smallest $Q$ estimation is an image with no ball rather than an image with a ball about
+to be missed by the agent: the DQN considers that, in terms of reward expectation, having missed a ball is worse than missing a ball. This
+reinforces the hypothesis that the DQN did not learn to play FIRE since not restarting the game means no reward.
 
@@
 \end{figure}
-
-Figure~\ref{fig:DDQN + DQNP} shows the evolution of reward per episode and of $Q$ estimates on 16,000,000 steps for the DDQN and 7,000,000 for the DQN with
-prioritized memory.
-
-One can observe that the DDQN seems to stabilize around an average reward per episode of 1.4 (though more training time would be needed to ensure convergence),
-like the DQN with uniform memory. Yet, $Q$ estimates of the DQN end up around $2$ whereas they are only half as big for the DDQN.
 
@@
 	\begin{subfigure}{.47\textwidth}
-		\vspace{-.15cm}
 		\includegraphics[width=\textwidth]{figures/dqn_prioritized_e_scores}
@@
 	\begin{subfigure}{.47\textwidth}
-		\vspace{-.15cm}
 		\includegraphics[width=\textwidth]{figures/dqn_prioritized_q_values}
@@
 
-\todo{Continuer (ici pour ne pas flinguer la mise en page) l'analyse comparative}
+Figure~\ref{fig:dqn uniform scores dist} shows the distribution of scores when letting the DQN play for 100 games. We can see that despite having an
+average score of 39.88 (with a standard deviation of 41.14), the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered
+enough for the DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
+and more importantly, the agent selects the action at each frame. This highest score of 219 is probably due to a non-anticipated breaking through
+the brick layer, leading to the ball destroying the upper layer whose bricks add 7 points to the score.
+
+Figure~\ref{fig:DDQN + DQNP} shows the evolution of reward per episode and of $Q$ estimates on 16,000,000 steps for the DDQN and 7,000,000 for the DQN with
+prioritized memory.
+
+One can observe that the DDQN seems to stabilize around an average reward per episode of 1.4 (though more training time would be needed to ensure convergence),
+like the DQN with uniform memory. Yet, $Q$ estimates of the DQN end up around $2$ whereas they are only half as big for the DDQN, which is the very point of
+the DDQN: avoid overoptimistic estimations~\citep{DBLP:journals/corr/WangFL15}.
+
+On the other hand, the addition of prioritized memory annihilates the results by making no score at all: Table~\ref{tab:scores breakout} shows that the performance
+is negative, so this agent is worse than random; it avoids the ball. Again, this is partly consistent with the results of~\cite{DBLP:journals/corr/SchaulQAS15}
+who also had worse performances on Breakout. Yet, the agent should still be able to learn how to play. This deficiency can also be due to the reduction of
+the replay memory size compared to the uniform memory or even to the instability of nonlinear estimators of the $Q$ function~\citep{tsitsiklis1996analysis}
+
+The distribution of scores per game for both DDQN and DQN with prioritized memory are shown in Figure~\ref{fig:scores DDQN + DQNP}.
+
+The distribution of the DDQN is very similar to the distribution of the DQN (if ignoring the score of 219 which must be the occurrence
+of a very unlikely event), only slightly shifted towards left, leading to the mode being [20, 30] for the DDQN versus [30, 40] for the DQN.
 
@@
 
-\todo{Les CNNs c'est instable, faut du temps et faire du café et bcp de parameter tuning -> solution plus stable ? Modèle linéaire (cf article d'origine) ?}
+The results obtained in this study don't pretend to outperform the original results by~\cite{mnih2015human} because the timing to develop and train the models,
+and the available computing power were restricted, causing the training to not be extensive and preventing from reproducing the learning with different
+parameters or even different initializations.
 
-\todo{Comme pour la prioritized faut mettre à jour les poids de tous les samples à chaque step, c'est compliqué faut optimiser ça}
+Furthermore, due to the computational cost of training a single model on different networks, we had to select a single game to test, out of the 49 games that were tested
+in the original work. The choice of \textit{Breakout} was finally made because of the performance obtained in the original work (over 10 times the score of the
+human player), as well as its simplicity in terms of features. Before making any attempt to improve on the described methods and models, one should first apply
+them to a wider variety of games, since the strength of this approach lies in its very general domain of application. Still, the efficiency of this method on
+different games, and then different types of gameplay, has been proved in~\cite{mnih2015human}.
 
-\todo{LSTM ?}
+When comparing scores obtained by the trained network and human players, it should be noted that even though the input available to both kind of players was the same,
+the learning process itself is not comparable. While human players learn in great part by applying knowledge from games they previously played, our agent had to
+infer every useful feature from scratch. It could be argued that each game taken individually was not designed to be the first and only game played by a human. In fact
+some elementary actions such as resuming a game immediately after a life loss, which would be performed instinctively by any human player, were not always performed by
+the model, resulting in many useless frames.
+
+Also, other improvements have been performed to this Deep Reinforcement Learning task since~\citep{mnih2015human}.
+
+Asynchronous advantage actor-critic (A3C) replaces the experience replay algorithm described hereabove with an asynchronous and parallel execution of multiple agents,
+each of whom evolves within its own, independent instance of the environment. The experience replay's role of decreasing the bias towards the most recent experience samples
+is performed by independently using different exploration policies and thus simultaneously experiencing a wide range of states. Moreover, replacing the experience replay
+algorithm additionally allows the use on-policy reinforcement learning methods such as actor-critic). The A3C algorithm also tackles the high hardware requirements of
+standard DQN algorithms, achieving better results in less training time (almost linear in the number of parallel agents) using a standard multi-core CPU~\citep{mnih2016asynchronous}
 
@@
 
-	\todo{Ajouter ordre de performance des 3 tests}
+We have implemented a standard DQN as presented in the original paper by Google, and we integrated two different improvements, namely a prioritized memory and a DDQN.
+As expected, both the DQN and the DDQN perform at least as well as Google's human tester, but the DQN with prioritized memory failed.
 
-\section{Acknowledgements}
+Yet, given the observed results and, the original work, our implementation seems to work, and could be trained on longer period to keep enhancing the results.
 
-  \todo{Remercier Lenaerts, Nowe, Google, OpenAI, Markov}
-
-%\newpage
+\newpage
 \footnotesize
